---
title: "UKHLS exploration (weighted kin) reftutorial 3.12"
format: html
editor: visual
---

## Reading and cleaning data

```{r}
#houskeeping
#Clear objects already in the environment – start with a clean slate
rm(list=ls())

#loading libraries
library(tidyverse)
library(svyVGAM)
library(sjlabelled)
library(desctable)
library(summarytools)
library(naniar)
library(survey)
library(svrep)
library(Hmisc)
library(srvyr)
library(marginaleffects)
library(haven)
library(catregs)
library(margins)
library(modelsummary)
library(zoo)
```

#### wave 1

```{r}
#wave 1 
inpath<-"D:/r git projects/ox-R/final essay/UKHLS n BHPS stata/UKDA-6614-stata/stata/stata13_se/"

a_indresp <- read_dta(file=paste0(inpath, "ukhls/a_indresp.dta")) %>%
 select(a_hidp, pidp, a_mastat_dv, a_strata, a_psu, a_hhorig, 
         a_indinus_xw,a_indpxus_xw, a_ind5mus_xw,a_ivfio, a_sampst,
         a_age_dv, a_sex_dv, a_hiqual_dv, a_paygu_dv, a_country,
         a_mastat_dv, a_racel_dv,a_fimnnet_dv,a_lvrel1,a_lvrel2,a_pns1pno,a_pns2pno,a_nchild_dv,a_nchresp,a_urban_dv,a_ohch16)


```

Some initial explorations:

```{r}
#wave 1: examine the data frame and variables of interest, including their distribution
view(dfSummary(a_indresp)) 
descr(a_indresp, headings = FALSE, stats = "common", transpose = TRUE)
names(a_indresp)
summary(a_indresp)
```

```{r}
# check the summary stats for the weights variables for different types of 
# respondents (first creating a smaller temporary df)

# Check: Are individual response weights 0 for proxy respondents?
a_indw <- a_indresp %>% 
  select(a_indinus_xw, a_indpxus_xw, a_ivfio, a_hhorig, a_sampst)

stby(data = a_indw, INDICES = a_indw$a_ivfio, 
     FUN = descr, stats = c("mean", "sd", "min", "max"),  
     transpose = TRUE)

# Check: Are all wave 1 weights zero for TSMs? NB. In wave 1, the only reason to# be a TSM is to be a non ethnic minority in the EMBS sample.
stby(data = a_indw, INDICES = a_indw$a_sampst, 
     FUN = descr, stats = c("mean", "sd", "min", "max"),  
     transpose = TRUE)
```

```{r}
rm(a_indw)
# Confirm that there are many PSUs and strata in the GB sample but only 1 strata in the Northern Ireland(NI, hhorig==3) sample 
tmp_df <- a_indresp %>% select(a_psu, a_strata, a_hhorig) 
stby(data = tmp_df, INDICES = tmp_df$a_hhorig, FUN = descr, 
     stats = c("mean", "sd", "min", "max"), transpose = TRUE)
```

#cleaning the data

```{r}
rm(tmp_df)
# replace missing values (integers -9 to -1) with NA
missval <- c(-9, -8, -7, -2, -1)
for (i in 1:5) {
  a_indresp <- a_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))}

#check NA
table(a_indresp$a_age_dv)
table(a_indresp$a_sex_dv)
a_indresp$a_sex_dv[a_indresp$a_sex_dv==0]<-NA
table(a_indresp$a_hiqual_dv)
table(a_indresp$a_paygu_dv<0)
table(a_indresp$a_country)
table(a_indresp$a_mastat_dv)
table(a_indresp$a_racel_dv)
table(a_indresp$a_fimnnet_dv<0)
a_indresp$a_fimnnet_dv[a_indresp$a_fimnnet_dv<0]<-NA
table(a_indresp$a_lvrel1)
table(a_indresp$a_lvrel2)
table(a_indresp$a_pns1pno)
table(a_indresp$a_pns2pno)
table(a_indresp$a_nchild_dv)
table(a_indresp$a_nchresp)
table(a_indresp$a_urban_dv)
table(a_indresp$a_ohch16)

rm(i,missval)
```

```{r}
#preparing variables for analysis (wave 1)
#num of parents (not censored)
a_indresp$parentnum<-a_indresp$a_lvrel1+a_indresp$a_lvrel2+a_indresp$a_pns1pno+a_indresp$a_pns2pno

#having one & two or more parents (censored)
a_indresp$parents<-c(0)
a_indresp$parents[a_indresp$parentnum==1]<-1
a_indresp$parents[a_indresp$parentnum>1]<-2

#living with parents or not (biological/step/adoptive)
a_indresp$lvpam<-c(0)
a_indresp$lvpam[a_indresp$a_pns1pno!=0]<-1
a_indresp$lvpaf<-c(0)
a_indresp$lvpaf[a_indresp$a_pns2pno!=0]<-1
a_indresp$lvpa<-a_indresp$lvpam+a_indresp$lvpaf
a_indresp<-a_indresp%>%select(-lvpam,-lvpaf)

#binary
a_indresp$lvpab<-c(0)
a_indresp$lvpab[a_indresp$lvpa!=0]<-1

#high ses and low ses (Degree/Other higher)
a_indresp$ses<-c(0)
a_indresp$ses[a_indresp$a_hiqual_dv%in%c(1,2)]<-1

#making nchild_dv a binary variable
a_indresp$child<-c(0)
a_indresp$child[a_indresp$a_nchild_dv>=1]<-1

#making partner a binary variable (Married/ In a registered same-sex civil partnership/Living as couple)
a_indresp$partner<-c(0)
a_indresp$partner[a_indresp$a_mastat_dv%in%c(2,3,10)]<-1

#makeing cohabitation a binary variable
a_indresp$cohab<-c(0)
a_indresp$cohab[a_indresp$a_mastat_dv%in%c(10)]<-1

#making married a binary variable
a_indresp$mar<-c(0)
a_indresp$mar[a_indresp$a_mastat_dv%in%c(2,3)]<-1

#logarithm of age
a_indresp$lnage<-log(a_indresp$a_age_dv)

#logarithm of income
a_indresp$lnincome<-log(a_indresp$a_fimnnet_dv)
#add 1 to all income values to avoid -inf in log transformation
a_indresp$incomeadd<-a_indresp$a_fimnnet_dv+1
a_indresp$lnincome<-log(a_indresp$incomeadd)
a_indresp<-a_indresp%>%select(-incomeadd)

```

```{r}
#preparing dependent variables for multinomial regression (wave 1) wech


```

```{r}
#preparing dependent variables for multinomial regression (wave 1) rout
##patterns of home-leaving 

a_indresp$homele<-c(0)

###living with parents(including those who have children)========================
a_indresp$homele[a_indresp$lvpa!=0]<-1

##route out 1: to live with a partner and children ==============================
###females
a_indresp$homele[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_sex_dv==2 & #females
                         a_indresp$a_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         a_indresp$a_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         a_indresp$a_nchild_dv!=0]<-2 #have a child under 16 in the hh


###males
a_indresp$homele[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_sex_dv==1 & #males
                         a_indresp$a_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         a_indresp$a_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         a_indresp$a_nchild_dv!=0]<-2#have a child under 16 in the hh



##route out 2: to live with a partner but no children=====================
###females
a_indresp$homele[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_sex_dv==2 & #females
                         a_indresp$a_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         a_indresp$a_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         a_indresp$a_nchild_dv==0]<-3 # don't have a child under 16 in the hh


###males
a_indresp$homele[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_sex_dv==1 & #males
                         a_indresp$a_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         a_indresp$a_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         a_indresp$a_nchild_dv==0]<-3 # don't have a child under 16 in the hh


##route out 3: forming a female-headed family============================
###females
a_indresp$homele[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_sex_dv==2 & #females
                         a_indresp$a_mastat_dv!=2 & #not married
                         a_indresp$a_mastat_dv!=3 & #not in civil partnership
                         a_indresp$a_mastat_dv!=10 & #not in cohabitation
                         a_indresp$a_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         a_indresp$a_nchild_dv!=0]<-4 #have a dependent child under 16 in the hh


###males
a_indresp$homele[a_indresp$lvpa==0 & # not living with parents
                         a_indresp$a_sex_dv==1 & #males
                         a_indresp$a_ohch16%in%c(1,2)]<-4 # (at least one) children under 16 not living in the same hh


##route out 4: to autonomous living =====================
###females
a_indresp$homele[a_indresp$lvpa==0 &
                         a_indresp$a_sex_dv==2 & #females
                         a_indresp$a_mastat_dv!=2 & #not married
                         a_indresp$a_mastat_dv!=3 & #not in civil partnership
                         a_indresp$a_mastat_dv!=10 & #not in cohabitation
                         a_indresp$a_mastat_dv%in%c(1,4,5,6,7)&
                         a_indresp$a_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         a_indresp$a_nchild_dv==0]<-5 # don't have a child under 16 in the hh

###males
a_indresp$homele[a_indresp$lvpa==0 &
                         a_indresp$a_sex_dv==1 & #males
                         a_indresp$a_mastat_dv!=2 & #not married
                         a_indresp$a_mastat_dv!=3 & #not in civil partnership
                         a_indresp$a_mastat_dv!=10 & #not in cohabitation
                         a_indresp$a_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         a_indresp$a_nchild_dv==0]<-5 # don't have a child under 16 in the hh


attr(a_indresp$homele,"labels")<-c(
"note: variables marked as missing/inapplicable/proxy/refusal/don't know are treated as not having dependent children under 16 outside the household",
"0:others(not living with parents: for females having at least one dependent child under 16 not living together; for males being single/separated from partners/divorced and living with at least one dependent child under 16; for both sexes if they have NA values in mastat_dv and nchild_dv);", "1:living with parents(including those who have children or partners);",
"2:route out 1: to live with a partner and children;",
"3:route out 2: to live with a partner but no children;",
"4:route out 3: forming a female-headed family",
"5:route out 4: to autonomous living")

```

#### wave 13

```{r}
#wave 13
m_indresp <- read_dta(file=paste0(inpath, "ukhls/m_indresp.dta")) %>%
 select(m_hidp, pidp, m_mastat_dv, m_strata, m_psu, m_hhorig, 
         m_indpxui_xw,m_indinui_xw,m_ivfio, m_sampst,
         m_age_dv, m_sex_dv, m_hiqual_dv, m_paygu_dv, m_country,
         m_mastat_dv, m_racel_dv,m_fimnnet_dv,m_lvrel1,m_lvrel2,m_lvrel9,m_lvrel10,m_pns1pno,m_pns2pno,m_nchild_dv,m_nchresp,m_urban_dv,m_ohch16)
```

```{r}
view(dfSummary(m_indresp)) 
descr(m_indresp, headings = FALSE, stats = "common", transpose = TRUE)
names(m_indresp)
summary(m_indresp)
```

```{r}
# Check: Are individual response weights 0 for proxy respondents?
m_indw <- m_indresp %>% 
  select(m_indpxui_xw,m_indinui_xw, m_ivfio, m_hhorig, m_sampst)

stby(data = m_indw, INDICES = m_indw$m_ivfio, 
     FUN = descr, stats = c("mean", "sd", "min", "max"),  
     transpose = TRUE)

rm(m_indw)

```

```{r}
# replace missing values (integers -9 to -1) with NA
missval <- c(-9, -8, -7, -2, -1)
for (i in 1:5) {
  m_indresp <- m_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))}

#check NA
table(m_indresp$m_age_dv)
table(m_indresp$m_sex_dv)
m_indresp$m_sex_dv[m_indresp$m_sex_dv==0]<-NA
table(m_indresp$m_hiqual_dv)
table(m_indresp$m_paygu_dv<0)
table(m_indresp$m_country)
table(m_indresp$m_mastat_dv)
table(m_indresp$m_racel_dv)
table(m_indresp$m_fimnnet_dv<0)
m_indresp$m_fimnnet_dv[m_indresp$m_fimnnet_dv<0]<-NA
table(m_indresp$m_lvrel1)
table(m_indresp$m_lvrel2)
table(m_indresp$m_lvrel9)
table(m_indresp$m_lvrel10)
table(m_indresp$m_pns1pno)
table(m_indresp$m_pns2pno)
table(m_indresp$m_nchild_dv)
table(m_indresp$m_nchresp)
table(m_indresp$m_urban_dv)
table(m_indresp$m_ohch16)

rm(i,missval)
```

```{r}
#preparing variables for analysis (wave 13)

#num of parents (not censored)
m_indresp$parentnum<-m_indresp$m_lvrel1+m_indresp$m_lvrel2+m_indresp$m_lvrel9+m_indresp$m_lvrel10+m_indresp$m_pns1pno+m_indresp$m_pns2pno

#having one & two or more parents (censored)
m_indresp$parents<-c(0)
m_indresp$parents[m_indresp$parentnum==1]<-1
m_indresp$parents[m_indresp$parentnum>1]<-2

#living with parents or not (biological/step/adoptive)
m_indresp$lvpam<-c(0)
m_indresp$lvpam[m_indresp$m_pns1pno!=0]<-1
m_indresp$lvpaf<-c(0)
m_indresp$lvpaf[m_indresp$m_pns2pno!=0]<-1
m_indresp$lvpa<-m_indresp$lvpam+m_indresp$lvpaf
m_indresp<-m_indresp%>%select(-lvpam,-lvpaf)
#binary
m_indresp$lvpab<-c(0)
m_indresp$lvpab[m_indresp$lvpa!=0]<-1


#high ses and low ses (Degree/Other higher)
m_indresp$ses<-c(0)
m_indresp$ses[m_indresp$m_hiqual_dv%in%c(1,2)]<-1

#making nchild_dv a binary variable
m_indresp$child<-c(0)
m_indresp$child[m_indresp$m_nchild_dv>=1]<-1

#making partner a binary variable (Married/ In a registered same-sex civil partnership/Living as couple)
m_indresp$partner<-c(0)
m_indresp$partner[m_indresp$m_mastat_dv%in%c(2,3,10)]<-1

#makeing cohabitation a binary variable
m_indresp$cohab<-c(0)
m_indresp$cohab[m_indresp$m_mastat_dv%in%c(10)]<-1

#making married a binary variable
m_indresp$mar<-c(0)
m_indresp$mar[m_indresp$m_mastat_dv%in%c(2,3)]<-1

#logarithm of age
m_indresp$lnage<-log(m_indresp$m_age_dv)

#logarithm of income
#add 1 to all income values to avoid -inf in log transformation
m_indresp$incomeadd<-m_indresp$m_fimnnet_dv+1
m_indresp$lnincome<-log(m_indresp$incomeadd)
m_indresp$lnincometest<-log(m_indresp$m_fimnnet_dv)
m_indresp<-m_indresp%>%select(-incomeadd,-lnincometest)
```

```{r}
#preparing dependent variables for multinomial regression (wave 13) wech


```

```{r}
#preparing dependent variables for multinomial regression (wave 13) rou
##patterns of home-leaving 
m_indresp$homele<-c(0)

###living with parents(including those who have children or partners)============
m_indresp$homele[m_indresp$lvpa!=0]<-1

##route out 1: to live with a partner and children ==============================
###females
m_indresp$homele[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_sex_dv==2 & #females
                         m_indresp$m_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         m_indresp$m_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         m_indresp$m_nchild_dv!=0]<-2 #have a child under 16 in the hh


###males
m_indresp$homele[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_sex_dv==1 & #males
                         m_indresp$m_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         m_indresp$m_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         m_indresp$m_nchild_dv!=0]<-2#have a child under 16 in the hh



##route out 2: to live with a partner but no children=====================
###females
m_indresp$homele[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_sex_dv==2 & #females
                         m_indresp$m_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         m_indresp$m_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         m_indresp$m_nchild_dv==0]<-3 # don't have a child under 16 in the hh


###males
m_indresp$homele[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_sex_dv==1 & #males
                         m_indresp$m_mastat_dv%in%c(2,3,10) & #married or in cohabitation
                         m_indresp$m_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         m_indresp$m_nchild_dv==0]<-3 # don't have a child under 16 in the hh


##route out 3: forming a female-headed family============================
###females
m_indresp$homele[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_sex_dv==2 & #females
                         m_indresp$m_mastat_dv!=2 & #not married
                         m_indresp$m_mastat_dv!=3 & #not in civil partnership
                         m_indresp$m_mastat_dv!=10 & #not in cohabitation
                         m_indresp$m_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         m_indresp$m_nchild_dv!=0]<-4 #have a dependent child under 16 in the hh


###males
m_indresp$homele[m_indresp$lvpa==0 & # not living with parents
                         m_indresp$m_sex_dv==1 & #males
                         m_indresp$m_ohch16%in%c(1,2)]<-4 # (at least one) children under 16 not living in the same hh


##route out 4: to autonomous living =====================
###females
m_indresp$homele[m_indresp$lvpa==0 &
                         m_indresp$m_sex_dv==2 & #females
                         m_indresp$m_mastat_dv!=2 & #not married
                         m_indresp$m_mastat_dv!=3 & #not in civil partnership
                         m_indresp$m_mastat_dv!=10 & #not in cohabitation
                         m_indresp$m_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         m_indresp$m_nchild_dv==0]<-5 # don't have a child under 16 in the hh

###males
m_indresp$homele[m_indresp$lvpa==0 &
                         m_indresp$m_sex_dv==1 & #males
                         m_indresp$m_mastat_dv!=2 & #not married
                         m_indresp$m_mastat_dv!=3 & #not in civil partnership
                         m_indresp$m_mastat_dv!=10 & #not in cohabitation
                         m_indresp$m_ohch16%in%c(3,NA) & # no child under 16 not living with the respondent
                         m_indresp$m_nchild_dv==0]<-5 # don't have a child under 16 in the hh

attr(m_indresp$homele,"labels")<-c(
"note: variables marked as missing/inapplicable/proxy/refusal/don't know are treated as not having dependent children under 16 outside the household",
"0:others(not living with parents: for females having at least one dependent child under 16 not living together; for males being single/separated from partners/divorced and living with at least one dependent child under 16; for both sexes if they have NA values in mastat_dv and nchild_dv);", "1:living with parents(including those who have children or partners);",
"2:route out 1: to live with a partner and children;",
"3:route out 2: to live with a partner but no children;",
"4:route out 3: forming a female-headed family",
"5:route out 4: to autonomous living")

```

```{r}
##checking for the variable homele
table(a_indresp$a_ohch16,a_indresp$a_sex_dv)
table(m_indresp$m_ohch16,m_indresp$m_sex_dv)

##temporary data================================================================
a_indrespy<-a_indresp[which(a_indresp$a_age_dv%in%c(25:50)),]
m_indrespy<-m_indresp[which(m_indresp$m_age_dv%in%c(25:50)),]

a_indrespyf<-a_indrespy[which(a_indrespy$a_sex_dv==2),]
m_indrespyf<-m_indrespy[which(m_indrespy$m_sex_dv==2),]

a_indrespym<-a_indrespy[which(a_indrespy$a_sex_dv==1),]
m_indrespym<-m_indrespy[which(m_indrespy$m_sex_dv==1),]

a_indrespyfc<-a_indrespyf[which(a_indrespyf$homele==0),]
m_indrespyfc<-m_indrespyf[which(m_indrespyf$homele==0),]

a_indrespymc<-a_indrespym[which(a_indrespym$homele==0),]
m_indrespymc<-m_indrespym[which(m_indrespym$homele==0),]

a_indrespyc<-a_indrespy[which(a_indrespym$homele==0),]
m_indrespyc<-m_indrespy[which(m_indrespym$homele==0),]

addmargins(table(a_indrespy$a_ohch16,a_indrespy$a_sex_dv))
addmargins(table(m_indrespy$m_ohch16,m_indrespy$m_sex_dv))

#for females not living with parents who have at least one dependent child under 16 outside the household (wave 1 n=211--215, wave 13 n=48--77)
addmargins(table(a_indrespyfc$a_ohch16,a_indrespyfc$a_mastat_dv,exclude="T"))
##females, marital status in c(1,4,5,6,7) and have no children are categorized as others
addmargins(table(m_indrespyfc$m_ohch16,m_indrespyfc$m_mastat_dv,exclude="T"))

#for males not living with parents who are single/separated from partners/divorced and have a dependent child under 16 in the household (wave 1 n=18--108, wave 13 n=7--52)
addmargins(table(a_indrespymc$a_ohch16,a_indrespymc$a_mastat_dv,exclude="T"))
addmargins(table(m_indrespymc$m_ohch16,m_indrespymc$m_mastat_dv,exclude="T"))

rm(a_indrespy,m_indrespy,a_indrespyf,m_indrespyf,a_indrespym,m_indrespym,a_indrespyfc,m_indrespyfc,a_indrespymc,m_indrespymc,a_indrespyc,m_indrespyc)

#checking how many female-headed families formed (by sex) in each wave
a_indresp%>%group_by(a_sex_dv)%>%filter(homele==3)%>%summarise(n=n())
m_indresp%>%group_by(m_sex_dv)%>%filter(homele==3)%>%summarise(n=n())
```

```{r}
#checking the marital status and whether having at least one dependent child for those who live with parents
a_indresp%>%group_by(a_sex_dv,a_nchild_dv)%>%filter(homele==1)%>%summarise(n=n())
a_indresp%>%group_by(a_sex_dv,a_mastat_dv)%>%filter(homele==1)%>%summarise(n=n())


```

### Adopt the weightings

```{r}
options(survey.lonely.psu="adjust")

svy_indresp1 <- svydesign(id=~a_psu, strata=~a_strata, 
                         weights=~a_indinus_xw, data=a_indresp)

svy_indresp2 <- svydesign(id=~m_psu, strata=~m_strata,weights=~m_indinui_xw, data=m_indresp)

#subset the data for respondents aged 25-50
svy_indresp1y<-subset(svy_indresp1, a_age_dv%in%c(25:50))

svy_indresp2y<-subset(svy_indresp2, m_age_dv%in%c(25:50))

```

```{r}
#estimating design effects
# calculate the mean pay for the sample, accounting for complex survey design
svymean(~a_paygu_dv,design=svy_indresp1,na.rm=TRUE,deff=TRUE)

#warning: deff not for the subsets of data
svymean(~a_paygu_dv,design=svy_indresp1y,na.rm=TRUE,deff=TRUE)

# compute DEFT: Remove # and replace "DEFF" in the command "sqrt(DEFF)" with value of DEFF computed in "svymean(~a_paygu_dv,svy_indresp,na.rm=TRUE,deff=TRUE)" 
#sqrt(DEFF)

#the Deff is 35.39. This means that a sample drawn using the current sampling plan needs to be 35.4 times the size needed if the sample was collected via an SRS
```

## Analyzing data (Cross-sectional)

### cross-sectional comparisons of two waves

Descriptive statistics (wave 1)

```{r}
#==================================================================================
#weighted number of children
svytable(~a_nchild_dv, design = svy_indresp1y)

#==================================================================================
#weighted sex 1:male, 2: female
svytable(~a_sex_dv, design = svy_indresp1y)
#The logit option fits a logistic regression model and computes a Wald-type interval on the log-odds scale, which is then transformed to the probability scale.
svyciprop(~I(a_sex_dv==2), svy_indresp1y, method="logit")

#weighted number of children by sex
svytable(~a_sex_dv+a_nchild_dv, design = svy_indresp1y)
#barplot for the average weighted number of children by sex
barplt<-svyby(~a_nchild_dv,~a_sex_dv,design=svy_indresp1y,na = TRUE, svymean)
barplot(barplt,beside=TRUE,legend=TRUE)
rm(barplt)
#a chi-squared test of weighted number of children by sex
svychisq(~a_sex_dv+a_nchild_dv, svy_indresp1y, statistic="adjWald")


#==================================================================================
#weighted education levels
svytable(~a_hiqual_dv, design = svy_indresp1y)
#weighted education levels by sex
svytable(~a_sex_dv+a_hiqual_dv, design = svy_indresp1y)
#a chi-squared test of weighted education levels by sex
svychisq(~a_sex_dv+a_hiqual_dv, svy_indresp1y, statistic="adjWald")

#==================================================================================
#weighted marital status
svytable(~a_mastat_dv, design = svy_indresp1y)
#weighted marital status by sex
svytable(~a_sex_dv+a_mastat_dv, design = svy_indresp1y)
#a chi-squared test of weighted marital status by sex
svychisq(~a_sex_dv+a_mastat_dv, svy_indresp1y, statistic="adjWald")

#==================================================================================
#weighted residential areas (rural/urban)  1:urban, 2:rural
svytable(~a_urban_dv, design = svy_indresp1y)
svyciprop(~I(a_urban_dv==2), svy_indresp1y, method="logit")


#==================================================================================
#weighted mean monthly income (after tax)
svymean(~a_fimnnet_dv,design=svy_indresp1y,na.rm=TRUE)
#??svyquantile(~a_fimnnet_dv, design = svy_indresp1y, na.rm=TRUE, c(.25,.5,.75),ci=TRUE)

#histogram
svyhist(~a_fimnnet_dv,design=svy_indresp1y,na.rm=TRUE)
#histogram of income distribution after log transformation
svyhist(~lnincome,design=svy_indresp1y,na.rm=TRUE)

#boxplots of income (including the plot for separate genders)
svyboxplot(~a_fimnnet_dv~1, svy_indresp1y, all.outliers=TRUE)
svyboxplot(~a_fimnnet_dv~factor(a_sex_dv), svy_indresp1y, all.outliers=TRUE)
#logged
svyboxplot(~lnincome~1, svy_indresp1y, all.outliers=TRUE)
svyboxplot(~lnincome~factor(a_sex_dv), svy_indresp1y, all.outliers=TRUE)

#==================================================================================
#weighted racial compositions
svytable(~a_racel_dv, design = svy_indresp1y)

#==================================================================================
#weighted age distribution
svymean(~a_age_dv,design=svy_indresp1y,na.rm=TRUE)
#histogram
svyhist(~a_age_dv,design=svy_indresp1y,na.rm=TRUE)
#boxplots of age (including the plot for separate genders)
svyboxplot(~a_age_dv~1, svy_indresp1y, all.outliers=TRUE)
svyboxplot(~a_age_dv~factor(a_sex_dv), svy_indresp1y, all.outliers=TRUE)

#==================================================================================
#weighted living with parents or not
svytable(~lvpab,design=svy_indresp1y)

#==================================================================================
#home-leaving patterns
svytable(~homele,design=svy_indresp1y)
```

Descriptive statistics (wave 13)

```{r}
#==================================================================================
#weighted number of children
svytable(~m_nchild_dv, design = svy_indresp2y)

#==================================================================================
#weighted sex 1:male, 2: female
svytable(~m_sex_dv, design = svy_indresp2y)
#The logit option fits a logistic regression model and computes a Wald-type interval on the log-odds scale, which is then transformed to the probability scale.
svyciprop(~I(m_sex_dv==2), svy_indresp2y, method="logit")

#weighted number of children by sex
svytable(~m_sex_dv+m_nchild_dv, design = svy_indresp2y)
#barplot for the average weighted number of children by sex
barplt<-svyby(~m_nchild_dv,~m_sex_dv,design=svy_indresp2y,na = TRUE, svymean)
barplot(barplt,beside=TRUE,legend=TRUE)
rm(barplt)
#a chi-squared test of weighted number of children by sex**
##???svychisq(~m_sex_dv+m_nchild_dv, svy_indresp2y, statistic="adjWald")Error in solve.default(V[use, use], Y[use]) : Lapack例行程序dgesv: 系统正好是奇异的: U[7,7] = 0

#==================================================================================
#weighted education levels
svytable(~m_hiqual_dv, design = svy_indresp2y)
#weighted education levels by sex
svytable(~m_sex_dv+m_hiqual_dv, design = svy_indresp2y)
#a chi-squared test of weighted education levels by sex
svychisq(~m_sex_dv+m_hiqual_dv, svy_indresp2y, statistic="adjWald")

#==================================================================================
#weighted marital status
svytable(~m_mastat_dv, design = svy_indresp2y)
#weighted marital status by sex
svytable(~m_sex_dv+m_mastat_dv, design = svy_indresp2y)
#a chi-squared test of weighted marital status by sex
svychisq(~m_sex_dv+m_mastat_dv, svy_indresp2y, statistic="adjWald")

#==================================================================================
#weighted residential areas (rural/urban)  1:urban, 2:rural
svytable(~m_urban_dv, design = svy_indresp2y)
svyciprop(~I(m_urban_dv==2), svy_indresp2y, method="logit")


#==================================================================================
#weighted mean monthly income (after tax)
svymean(~m_fimnnet_dv,design=svy_indresp2y,na.rm=TRUE)
#??svyquantile(~m_fimnnet_dv, design = svy_indresp2y, na.rm=TRUE, c(.25,.5,.75),ci=TRUE)Can't merge the outer name `ci` with a vector of length > 1.Please supply a `.name_spec` specification. 

#histogram
svyhist(~m_fimnnet_dv,design=svy_indresp2y,na.rm=TRUE)
#histogram of income distribution after log transformation
svyhist(~lnincome,design=svy_indresp2y,na.rm=TRUE)

#boxplots of income (including the plot for separate genders)
svyboxplot(~m_fimnnet_dv~1, svy_indresp2y, all.outliers=TRUE)
svyboxplot(~m_fimnnet_dv~factor(m_sex_dv), svy_indresp2y, all.outliers=TRUE)
#logged
svyboxplot(~lnincome~1, svy_indresp2y, all.outliers=TRUE)
svyboxplot(~lnincome~factor(m_sex_dv), svy_indresp2y, all.outliers=TRUE)
#==================================================================================
#weighted racial compositions
svytable(~m_racel_dv, design = svy_indresp2y)

#==================================================================================
#weighted age distribution
svymean(~m_age_dv,design=svy_indresp2y,na.rm=TRUE)
#histogram
svyhist(~m_age_dv,design=svy_indresp2y,na.rm=TRUE)
#boxplots of age (including the plot for separate genders)
svyboxplot(~m_age_dv~1, svy_indresp2y, all.outliers=TRUE)
svyboxplot(~m_age_dv~factor(m_sex_dv), svy_indresp2y, all.outliers=TRUE)

#==================================================================================
#weighted living with parents or not
svytable(~lvpab,design=svy_indresp2y)

#==================================================================================
#home-leaving patterns
svytable(~homele,design=svy_indresp2y)
```

##### Model 1

```{r}
m1a<-svyglm(parents~ses+a_age_dv+a_sex_dv+lnincome+factor(a_racel_dv),design=svy_indresp1y,na.action=na.omit,family=poisson())
        
m1b<-svyglm(parents~ses+m_age_dv+m_sex_dv+lnincome+factor(m_racel_dv),design=svy_indresp2y,na.action=na.omit,family=poisson())

summary(m1a)
summary(m1b)

#grandparents effects(for those who have children under 16 living together)

```

##### Model 2

```{r}
#whether married or not
m2ma<-svyglm(mar~ses+a_age_dv+a_sex_dv+lnincome+factor(a_racel_dv),design=svy_indresp1y,na.action=na.omit,family=quasibinomial)
m2mb<-

#whether cohabiting with others or not
m2ca<-
m2cb<-

```

##### Model 3

```{r}
#whether having at least one dependent child under 16
```

##### Model 4

```{r}

```

##### Model 5

```{r}
#m5a<-svy_vglm(homele~ses+a_age_dv+a_sex_dv+lnincome+factor(a_racel_dv)+a_urban_dv+parents,family=multinomial(refLevel=1),design=svy_indresp1y,na.action=na.omit)

#mult_eth<- svy_vglm(RIDRETH1~RIDAGEYR+DMDEDUC,family=multinomial(refLevel=3), design=nhdes)

#If you are noticing lag in the RStudio console, i.e. slow responsiveness, you may need to clear the console cache. Try clearing the console cache with ctrl+L and check if that improves performance.

m5a<-svy_vglm(homele~ses+a_age_dv+a_sex_dv+lnincome+factor(a_racel_dv)+a_urban_dv+parents,family=multinomial(refLevel=1),design=svy_indresp1y,na.action=na.omit)
summary(m5a)
```

### amatching parents and their children for analysis

### matching couples for analysis

### finding out who are the leavers or stayers

## Analyzing data (Longitudinal)
