---
title: "UKHLS exploration (weighted kin) 4.6 (2)"
format: html
editor: visual
---

## Reading and cleaning data

```{r}
#houskeeping
#Clear objects already in the environment – start with a clean slate
rm(list=ls())

#loading libraries
library(tidyverse)
library(svyVGAM)
library(sjlabelled)
library(desctable)
library(summarytools)
library(naniar)
library(survey)
library(svrep)
library(Hmisc)
library(srvyr)
library(marginaleffects)
library(haven)
library(catregs)
library(margins)
library(modelsummary)
library(zoo)
library(mice)
library(stargazer)
library(texreg)
library(VIM)
library(lattice)
library(ggplot2)
```

### Cross-wave dat

```{r}
inpath<-"D:/r git projects/ox-R/final essay/UKHLS n BHPS stata/UKDA-6614-stata/stata/stata13_se/"
xwavedat<-read_dta(file=paste0(inpath, "ukhls/xwavedat.dta")) %>%
 dplyr::select(pidp,maju,paju,maedqf,paedqf)
```

```{r}
missval <- c(-9, -8, -7, -2, -1)
for (i in 1:5) {
  xwavedat<- xwavedat %>%
    mutate_all(., list(~na_if(., missval[i])))
  
}
```

### Wave1

```{r}
#wave 1 
a_indresp <- read_dta(file=paste0(inpath, "ukhls/a_indresp.dta")) %>%
 dplyr::select(pidp, a_mastat_dv, a_strata, a_psu,
         a_indinus_xw,
         a_age_dv, a_sex_dv, a_hiqual_dv,
         a_mastat_dv, a_racel_dv,a_fimnnet_dv,a_lvrel1,a_lvrel2,a_pns1pno,a_pns2pno,a_nchild_dv,a_urban_dv,a_panssec8_dv,a_manssec8_dv)


a_indrespj<-a_indresp%>%dplyr::select(pidp,a_panssec8_dv,a_manssec8_dv)

library(plyr)
a_indresp<-join_all(list(a_indresp,xwavedat), by='pidp', type='left')
```

```{r}
#rename pa ses, job and edu columns
a_indresp <-a_indresp%>% 
  dplyr::rename(
   pases=a_panssec8_dv,
    mases=a_manssec8_dv
    )
```

### Wave 2\~12

```{r}
#wave 2
b_indresp <- read_dta(file=paste0(inpath, "ukhls/b_indresp.dta")) %>%
 dplyr::select(pidp,b_panssec8_dv,b_manssec8_dv)

#wave 3
c_indresp <- read_dta(file=paste0(inpath, "ukhls/c_indresp.dta")) %>%
 dplyr::select(pidp,c_panssec8_dv,c_manssec8_dv)

#wave 4
d_indresp <- read_dta(file=paste0(inpath, "ukhls/d_indresp.dta")) %>%
 dplyr::select(pidp,d_panssec8_dv,d_manssec8_dv)

#wave 5
e_indresp <- read_dta(file=paste0(inpath, "ukhls/e_indresp.dta")) %>%
 dplyr::select(pidp,e_panssec8_dv,e_manssec8_dv)

#wave 6
f_indresp <- read_dta(file=paste0(inpath, "ukhls/f_indresp.dta")) %>%
 dplyr::select(pidp,f_panssec8_dv,f_manssec8_dv)

#wave 7
g_indresp <- read_dta(file=paste0(inpath, "ukhls/g_indresp.dta")) %>%
 dplyr::select(pidp,g_panssec8_dv,g_manssec8_dv)

#wave 8
h_indresp <- read_dta(file=paste0(inpath, "ukhls/h_indresp.dta")) %>%
 dplyr::select(pidp,h_panssec8_dv,h_manssec8_dv)

#wave 9
i_indresp <- read_dta(file=paste0(inpath, "ukhls/i_indresp.dta")) %>%
 dplyr::select(pidp,i_panssec8_dv,i_manssec8_dv)

#wave 10
j_indresp <- read_dta(file=paste0(inpath, "ukhls/j_indresp.dta")) %>%
 dplyr::select(pidp,j_panssec8_dv,j_manssec8_dv)

#wave 11
k_indresp <- read_dta(file=paste0(inpath, "ukhls/k_indresp.dta")) %>%
 dplyr::select(pidp,k_panssec8_dv,k_manssec8_dv)

#wave 12
l_indresp <- read_dta(file=paste0(inpath, "ukhls/l_indresp.dta")) %>%
 dplyr::select(pidp,l_panssec8_dv,l_manssec8_dv)
```

### Wave13

```{r}
#wave 13
m_indresp <- read_dta(file=paste0(inpath, "ukhls/m_indresp.dta")) %>%
 dplyr::select(pidp, m_mastat_dv, m_strata, m_psu, 
         m_indpxui_xw,
         m_age_dv, m_sex_dv, m_hiqual_dv, 
         m_mastat_dv, m_racel_dv,m_fimnnet_dv,m_lvrel1,m_lvrel2,m_lvrel9,m_lvrel10,m_pns1pno,m_pns2pno,m_nchild_dv,m_urban_dv,m_panssec8_dv,m_manssec8_dv)

```

```{r}
# recode missing values for all variables using for loop and mutate_all
for (i in 1:5) {
  a_indresp<- a_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}

for (i in 1:5) {
  a_indrespj<- a_indrespj %>%
    mutate_all(., list(~na_if(., missval[i])))
}

for (i in 1:5) {
  b_indresp<- b_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  c_indresp<- c_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  d_indresp<- d_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  e_indresp<- e_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  f_indresp<- f_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  g_indresp<- g_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  h_indresp<- h_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  i_indresp<- i_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  j_indresp<- j_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  k_indresp<- k_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  l_indresp<- l_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  m_indresp<- m_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
  
}
```

```{r}
#wave 13 join
library(plyr)
m_indresp<-join_all(list(m_indresp,a_indrespj,b_indresp,c_indresp,d_indresp,e_indresp,f_indresp,g_indresp,h_indresp,i_indresp,j_indresp,k_indresp,l_indresp,xwavedat), by='pidp', type='left')

m_indresp<-m_indresp%>% mutate(
                           mases=coalesce(a_manssec8_dv,b_manssec8_dv,c_manssec8_dv,d_manssec8_dv,e_manssec8_dv,f_manssec8_dv,g_manssec8_dv,h_manssec8_dv,i_manssec8_dv,j_manssec8_dv,k_manssec8_dv,l_manssec8_dv,m_manssec8_dv),
                           pases=coalesce(a_panssec8_dv,b_panssec8_dv,c_panssec8_dv,d_panssec8_dv,e_panssec8_dv,f_panssec8_dv,g_panssec8_dv,h_panssec8_dv,i_panssec8_dv,j_panssec8_dv,k_panssec8_dv,l_panssec8_dv,m_panssec8_dv)) 
```

```{r}
rm(missval,a_indrespj,b_indresp,c_indresp,d_indresp,e_indresp,f_indresp,g_indresp,h_indresp,i_indresp,j_indresp,k_indresp,l_indresp,xwavedat)
```

# Cleaning data

### Wave 1

```{r}
a_indresp$a_sex_dv[a_indresp$a_sex_dv==0]<-NA
a_indresp$a_fimnnet_dv[a_indresp$a_fimnnet_dv<0]<-NA
```

```{r}
#preparing variables for analysis (wave 1)
#num of parents
a_indresp$lvpam<-c(0)
a_indresp$lvpam[a_indresp$a_pns1pno!=0]<-1
a_indresp$lvpaf<-c(0)
a_indresp$lvpaf[a_indresp$a_pns2pno!=0]<-1
a_indresp$parentnum<-a_indresp$a_lvrel1+a_indresp$a_lvrel2+a_indresp$lvpam+a_indresp$lvpaf


#living with parents or not (biological/step/adoptive)
a_indresp$lvpa<-a_indresp$lvpam+a_indresp$lvpaf
a_indresp<-a_indresp%>%dplyr::select(-lvpam,-lvpaf)

#binary
a_indresp$lvpab<-c(0)
a_indresp$lvpab[a_indresp$lvpa!=0]<-1

#level of education completed
a_indresp$edu<-c(0)
a_indresp$edu[a_indresp$a_hiqual_dv%in%c(9)]<-1 #low (everything below)
a_indresp$edu[a_indresp$a_hiqual_dv%in%c(3,4,5)]<-2 #middle (completed A-level or secondary high school)
a_indresp$edu[a_indresp$a_hiqual_dv%in%c(1,2)]<-3 #high (have a degree)
a_indresp$edu[a_indresp$edu==0]<-NA

print(attr(a_indresp$a_hiqual_dv,"labels"))
a_indresp$edu<-factor(a_indresp$edu,levels=c(1,2,3),labels=c('low',"middle","high"))


#making nchild_dv a binary variable
a_indresp$child<-c(0)
a_indresp$child[a_indresp$a_nchild_dv>=1]<-1

#age group
a_indresp$age_group<-c(0)
a_indresp$age_group[a_indresp$a_age_dv%in%c(16:20)]<-1
a_indresp$age_group[a_indresp$a_age_dv%in%c(21:25)]<-2
a_indresp$age_group[a_indresp$a_age_dv%in%c(26:30)]<-3
a_indresp$age_group[a_indresp$a_age_dv%in%c(31:35)]<-4

a_indresp$age_group<-factor(a_indresp$age_group,levels=c(1,2,3,4),labels=c("16-20","21-25","26-30","31-35"))

#logarithm of income
a_indresp$lnincome<-log(a_indresp$a_fimnnet_dv)
#add 1 to all income values to avoid -inf in log transformation
a_indresp$incomeadd<-a_indresp$a_fimnnet_dv+1
a_indresp$lnincome<-log(a_indresp$incomeadd)
a_indresp<-a_indresp%>%dplyr::select(-incomeadd)

#racel groups 
a_indresp$racel<-c(0)
a_indresp$racel[a_indresp$a_racel_dv%in%c(1,2,3,4)]<-1 #Whites
a_indresp$racel[a_indresp$a_racel_dv%in%c(14,15)]<-2 #African Caribbean
a_indresp$racel[a_indresp$a_racel_dv%in%c(9,10,11)]<-3 #Indiani,Pakistani, Bangladeshi
a_indresp$racel[a_indresp$a_racel_dv%in%c(12,13)]<-4 #Other Asian
a_indresp$racel[a_indresp$racel==0]<-5 #Other

a_indresp$racel<-factor(a_indresp$racel,levels=c(1,2,3,4,5),labels=c("Whites","African Caribbean","Indiani,Pakistani, Bangladeshi","Other Asian","Other"))
```

```{r}
#father ses
a_indresp$fases<-c(0)
a_indresp$fases[a_indresp$pases==8]<-1 #Routine 
a_indresp$fases[a_indresp$pases==7]<-2 #Semi-routine 
a_indresp$fases[a_indresp$pases==6]<-3 #Lower supervisory & technical
a_indresp$fases[a_indresp$pases==5]<-4 #Small employers & own account
a_indresp$fases[a_indresp$pases==4]<-5 #Intermediate 
a_indresp$fases[a_indresp$pases==3]<-6 #Lower management & professional
a_indresp$fases[a_indresp$pases==2]<-7 #Higher professional
a_indresp$fases[a_indresp$pases==1]<-8 #Large employers & higher management 

#dummy variable for pases missing
#a_indresp$fasesna<-ifelse(a_indresp$fases==0, 1, 0)
a_indresp$fases[a_indresp$pases==0]<-NA

a_indresp$fasesXage<-a_indresp$fases*a_indresp$a_age_dv

#for parental education
#father
a_indresp$paedu<-c(0)
a_indresp$paedu[a_indresp$paedqf%in%c(1,2,97)]<-1 #low (everything below, the category "other" is included)
a_indresp$paedu[a_indresp$paedqf%in%c(3,4)]<-2 #middle (complete A-level or have some qualifications)
a_indresp$paedu[a_indresp$paedqf%in%c(5)]<-3 #high (have a degree)
a_indresp$paedu[a_indresp$paedu==0]<-NA

a_indresp$paedu<-factor(a_indresp$paedu,levels=c(1,2,3),labels=c('low',"middle","high"))

#mother
a_indresp$maedu<-c(0)
a_indresp$maedu[a_indresp$maedqf%in%c(1,2,97)]<-1 #low (everything below, the category "other" is included)
a_indresp$maedu[a_indresp$maedqf%in%c(3,4)]<-2 #middle (complete A-level or have some qualifications)
a_indresp$maedu[a_indresp$maedqf%in%c(5)]<-3 #high (have a degree)
a_indresp$maedu[a_indresp$maedu==0]<-NA

a_indresp$maedu<-factor(a_indresp$maedu,levels=c(1,2,3),labels=c('low',"middle","high"))

#for whether mother working at age 14
a_indresp$mawork<-c(0)
a_indresp$mawork[a_indresp$maju==2]<-1 #notworking
a_indresp$mawork[a_indresp$maju==1]<-2 #working
a_indresp$mawork[a_indresp$maju%in%c(3,4)]<-3 #Mother deceased/Mother not living with respondent so don't know
a_indresp$mawork[a_indresp$mawork==0]<-NA #missing

a_indresp$mawork<-factor(a_indresp$mawork,levels=c(1,2,3),labels=c("notworking","working","Mother deceased/not living with respondent"))

```

```{r}
#preparing dependent variables for multinomial regression (wave 1)
##patterns of home-leaving
###living with parents(including those who have children)========================
a_indresp$leave[a_indresp$lvpa!=0]<-1
##route out 1: to live with a partner ==============================
a_indresp$leave[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_mastat_dv%in%c(2,3,10)  #married 
                         ]<-2 

##route out 2: to live without a partner=====================
a_indresp$leave[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_mastat_dv%in%c(0,1,4,5,6,7,8,9,10)  
                         ]<-3 
```

```{r}
#preparing dependent variables for multinomial regression (wave 1) 
##patterns of home-leaving (distinguish between marriage and cohabitation) 
a_indresp$ledes<-c(0)

###living with parents(including those who have children)========================
a_indresp$ledes[a_indresp$lvpa!=0]<-1

##route out 1: to live with a partner(marriage) ==============================
a_indresp$ledes[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_mastat_dv%in%c(2,3)  #married 
                         ]<-2 

##route out 2: to live with a partner(cohabitation) ==============================
a_indresp$ledes[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_mastat_dv%in%c(10)  #in cohabitation
                         ]<-3 

##route out 3: to live without a partner=====================
a_indresp$ledes[a_indresp$lvpa==0 & #not living with parents
                         a_indresp$a_mastat_dv%in%c(0,1,4,5,6,7,8,9)  #others
                         ]<-4 



attr(a_indresp$ledes,"labels")<-c(
"NA:n=13,having missing values in marital status;", 
"1:living with parents;",
"2:route out 1:leaving with partner (marriage) ;",
"3:route out 2:leaving with partner (cohabitation);",
"4:route out 3:leaving without a partner")


a_indresp$ledes<-factor(a_indresp$ledes,levels=c(1,2,3,4),labels=c("living with parents","leaving with partner (marriage)","leaving with partner (cohabitation)","leaving without a partner"))
```

### Wave 13

```{r}
m_indresp<-m_indresp%>%dplyr::select(pidp, m_mastat_dv, m_strata, m_psu, 
         m_indpxui_xw,
         m_age_dv, m_sex_dv, m_hiqual_dv, 
         m_mastat_dv, m_racel_dv,m_fimnnet_dv,m_lvrel1,m_lvrel2,m_lvrel9,m_lvrel10,m_pns1pno,m_pns2pno,m_nchild_dv,m_urban_dv,maju,paju,maedqf,paedqf,mases,pases)
```

```{r}
m_indresp$m_sex_dv[m_indresp$m_sex_dv==0]<-NA
m_indresp$m_fimnnet_dv[m_indresp$m_fimnnet_dv<0]<-NA
```

```{r}
#preparing variables for analysis (wave 13)
#num of parents
m_indresp$lvpam<-c(0)
m_indresp$lvpam[m_indresp$m_pns1pno!=0]<-1
m_indresp$lvpaf<-c(0)
m_indresp$lvpaf[m_indresp$m_pns2pno!=0]<-1
m_indresp$parentnum<-m_indresp$m_lvrel1+m_indresp$m_lvrel2+m_indresp$lvpam+m_indresp$lvpaf


#living with parents or not (biological/step/adoptive)
m_indresp$lvpa<-m_indresp$lvpam+m_indresp$lvpaf
m_indresp<-m_indresp%>%dplyr::select(-lvpam,-lvpaf)

#binary
m_indresp$lvpab<-c(0)
m_indresp$lvpab[m_indresp$lvpa!=0]<-1

#level of education completed
m_indresp$edu<-c(0)
m_indresp$edu[m_indresp$m_hiqual_dv%in%c(9)]<-1 #low (everything below)
m_indresp$edu[m_indresp$m_hiqual_dv%in%c(3,4,5)]<-2 #middle (completed A-level or secondary high school)
m_indresp$edu[m_indresp$m_hiqual_dv%in%c(1,2)]<-3 #high (have a degree)
m_indresp$edu[m_indresp$edu==0]<-NA

print(attr(m_indresp$m_hiqual_dv,"labels"))
m_indresp$edu<-factor(m_indresp$edu,levels=c(1,2,3),labels=c('low',"middle","high"))


#making nchild_dv a binary variable
m_indresp$child<-c(0)
m_indresp$child[m_indresp$m_nchild_dv>=1]<-1

#age group
m_indresp$age_group<-c(0)
m_indresp$age_group[m_indresp$m_age_dv%in%c(16:20)]<-1
m_indresp$age_group[m_indresp$m_age_dv%in%c(21:25)]<-2
m_indresp$age_group[m_indresp$m_age_dv%in%c(26:30)]<-3
m_indresp$age_group[m_indresp$m_age_dv%in%c(31:35)]<-4

m_indresp$age_group<-factor(m_indresp$age_group,levels=c(1,2,3,4),labels=c("16-20","21-25","26-30","31-35"))

#logarithm of income
m_indresp$lnincome<-log(m_indresp$m_fimnnet_dv)
#add 1 to all income values to avoid -inf in log transformation
m_indresp$incomeadd<-m_indresp$m_fimnnet_dv+1
m_indresp$lnincome<-log(m_indresp$incomeadd)
m_indresp<-m_indresp%>%dplyr::select(-incomeadd)

#racel groups 
m_indresp$racel<-c(0)
m_indresp$racel[m_indresp$m_racel_dv%in%c(1,2,3,4)]<-1 #Whites
m_indresp$racel[m_indresp$m_racel_dv%in%c(14,15)]<-2 #African Caribbean
m_indresp$racel[m_indresp$m_racel_dv%in%c(9,10,11)]<-3 #Indiani,Pakistani, Bangladeshi
m_indresp$racel[m_indresp$m_racel_dv%in%c(12,13)]<-4 #Other Asian
m_indresp$racel[m_indresp$racel==0]<-5 #Other

m_indresp$racel<-factor(m_indresp$racel,levels=c(1,2,3,4,5),labels=c("Whites","African Caribbean","Indiani,Pakistani, Bangladeshi","Other Asian","Other"))

#
```

```{r}
#father ses
m_indresp$fases<-c(0)
m_indresp$fases[m_indresp$pases==8]<-1 #Routine 
m_indresp$fases[m_indresp$pases==7]<-2 #Semi-routine 
m_indresp$fases[m_indresp$pases==6]<-3 #Lower supervisory & technical
m_indresp$fases[m_indresp$pases==5]<-4 #Small employers & own account
m_indresp$fases[m_indresp$pases==4]<-5 #Intermediate 
m_indresp$fases[m_indresp$pases==3]<-6 #Lower management & professional
m_indresp$fases[m_indresp$pases==2]<-7 #Higher professional
m_indresp$fases[m_indresp$pases==1]<-8 #Large employers & higher management 

#dummy variable for pases missing
#m_indresp$fasesna<-ifelse(m_indresp$fases==0, 1, 0)
m_indresp$fases[m_indresp$pases==0]<-NA

m_indresp$fasesXage<-m_indresp$fases*m_indresp$m_age_dv

#for parental education
#father
m_indresp$paedu<-c(0)
m_indresp$paedu[m_indresp$paedqf%in%c(1,2,97)]<-1 #low (everything below, the category "other" is included)
m_indresp$paedu[m_indresp$paedqf%in%c(3,4)]<-2 #middle (complete A-level or have some qualifications)
m_indresp$paedu[m_indresp$paedqf%in%c(5)]<-3 #high (have a degree)
m_indresp$paedu[m_indresp$paedu==0]<-NA

m_indresp$paedu<-factor(m_indresp$paedu,levels=c(1,2,3),labels=c('low',"middle","high"))

#mother
m_indresp$maedu<-c(0)
m_indresp$maedu[m_indresp$maedqf%in%c(1,2,97)]<-1 #low (everything below, the category "other" is included)
m_indresp$maedu[m_indresp$maedqf%in%c(3,4)]<-2 #middle (complete A-level or have some qualifications)
m_indresp$maedu[m_indresp$maedqf%in%c(5)]<-3 #high (have a degree)
m_indresp$maedu[m_indresp$maedu==0]<-NA

m_indresp$maedu<-factor(m_indresp$maedu,levels=c(1,2,3),labels=c('low',"middle","high"))

#for whether mother working at age 14
m_indresp$mawork<-c(0)
m_indresp$mawork[m_indresp$maju==2]<-1 #notworking
m_indresp$mawork[m_indresp$maju==1]<-2 #working
m_indresp$mawork[m_indresp$maju%in%c(3,4)]<-3 #Mother deceased/Mother not living with respondent so don't know
m_indresp$mawork[m_indresp$mawork==0]<-NA #missing

m_indresp$mawork<-factor(m_indresp$mawork,levels=c(1,2,3),labels=c("notworking","working","Mother deceased/not living with respondent"))

```

```{r}
#preparing dependent variables for multinomial regression (wave 1)
##patterns of home-leaving
###living with parents(including those who have children)========================
m_indresp$leave[m_indresp$lvpa!=0]<-1
##route out 1: to live with a partner ==============================
m_indresp$leave[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_mastat_dv%in%c(2,3,10)  #married 
                         ]<-2 

##route out 2: to live without a partner=====================
m_indresp$leave[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_mastat_dv%in%c(0,1,4,5,6,7,8,9,10)  
                         ]<-3 
```

```{r}
#preparing dependent variables for multinomial regression (wave 13) 
##patterns of home-leaving 
m_indresp$ledes<-c(0)

###living with parents(including those who have children)========================
m_indresp$ledes[m_indresp$lvpa!=0]<-1

##route out 1: to live with a partner(marriage) ==============================
m_indresp$ledes[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_mastat_dv%in%c(2,3)  #married
                         ]<-2 

##route out 2: to live with a partner(cohabitation) ==============================
m_indresp$ledes[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_mastat_dv%in%c(10)  #in cohabitation
                         ]<-3 

##route out 3: to live without a partner=====================
###females
m_indresp$ledes[m_indresp$lvpa==0 & #not living with parents
                         m_indresp$m_mastat_dv%in%c(0,1,4,5,6,7,8,9)  #others
                         ]<-4 


attr(m_indresp$ledes,"labels")<-c(
"NA:n=80,having missing values in marital status;", 
"1:living with parents;",
"2:route out 1:leaving with partner (marriage) ;",
"3:route out 2:leaving with partner (cohabitation);",
"4:route out 3:leaving without a partner")


m_indresp$ledes<-factor(m_indresp$ledes,levels=c(1,2,3,4),labels=c("living with parents","leaving with partner (marriage)","leaving with partner (cohabitation)","leaving without a partner"))
```

## Adopt the weightings

```{r}
options(survey.lonely.psu="adjust")

svy_indresp1 <- svydesign(id=~a_psu, strata=~a_strata, 
                         weights=~a_indinus_xw, data=a_indresp)

svy_indresp2 <- svydesign(id=~m_psu, strata=~m_strata,weights=~m_indpxui_xw, data=m_indresp)

#subset the data for respondents aged 25-50
svy_indresp1y<-subset(svy_indresp1, a_age_dv%in%c(16:35))

svy_indresp2y<-subset(svy_indresp2, m_age_dv%in%c(16:35))


#for males and females
#females
svy_indresp1yf<-subset(svy_indresp1y, a_sex_dv%in%c(2))
svy_indresp2yf<-subset(svy_indresp2y, m_sex_dv%in%c(2))


#males
svy_indresp1ym<-subset(svy_indresp1y, a_sex_dv%in%c(1))
svy_indresp2ym<-subset(svy_indresp2y, m_sex_dv%in%c(1))


```

## Model building with imputations

### dealing with haven issues

```{r}
a_indrespt<-a_indresp
a_indrespty<-a_indresp[which(a_indresp$a_age_dv%in%c(16:35)),]
a_indresptf<-a_indrespty[which(a_indrespty$a_sex_dv%in%c(2)),]

m_indrespt<-m_indresp
m_indrespty<-m_indresp[which(m_indresp$m_age_dv%in%c(16:35)),]
m_indresptf<-m_indrespty[which(m_indrespty$m_sex_dv%in%c(2)),]
```

```{r}
#wave 1
a_indrespt<-a_indrespt%>%dplyr::select(a_indinus_xw,a_strata,a_psu,a_mastat_dv,a_sex_dv,a_nchild_dv,a_urban_dv,mases,lvpab,edu,age_group,lnincome,racel,fases,fasesXage,paedu,maedu,mawork,leave,ledes)

#as numeric
a_indrespt$a_indinus_xw<-haven::zap_labels(a_indrespt$a_indinus_xw)
a_indrespt$lnincome<-haven::zap_labels(a_indrespt$lnincome)
a_indrespt$fasesXage<-haven::zap_labels(a_indrespt$fasesXage)
a_indrespt$fases<-haven::zap_labels(a_indrespt$fases)
a_indrespt$mases<-haven::zap_labels(a_indrespt$mases)

#as factor
a_indrespt$a_mastat_dv<-haven::as_factor(a_indrespt$a_mastat_dv)
a_indrespt$a_strata<-haven::as_factor(a_indrespt$a_strata)
a_indrespt$a_psu<-haven::as_factor(a_indrespt$a_psu)
a_indrespt$a_sex_dv<-haven::as_factor(a_indrespt$a_sex_dv)
a_indrespt$a_nchild_dv<-haven::as_factor(a_indrespt$a_nchild_dv)
a_indrespt$a_urban_dv<-haven::as_factor(a_indrespt$a_urban_dv)
a_indrespt$lvpab<-haven::as_factor(a_indrespt$lvpab)
a_indrespt$racel<-haven::as_factor(a_indrespt$racel)
a_indrespt$ledes<-haven::as_factor(a_indrespt$ledes)
a_indrespt$mawork <-haven::as_factor(a_indrespt$mawork)
a_indrespt$edu<-haven::as_factor(a_indrespt$edu)
a_indrespt$age_group<-haven::as_factor(a_indrespt$age_group)
a_indrespt$paedu <-haven::as_factor(a_indrespt$paedu )
a_indrespt$maedu  <-haven::as_factor(a_indrespt$maedu )


#wave 13 a_indinus_xw
m_indrespt<-m_indrespt%>%dplyr::select(m_indpxui_xw,m_strata,m_psu,m_mastat_dv,m_sex_dv,m_nchild_dv,m_urban_dv,mases,lvpab,edu,age_group,lnincome,racel,fases,fasesXage,paedu,maedu,mawork,leave,ledes)

#as numeric
m_indrespt$m_indpxui_xw<-haven::zap_labels(m_indrespt$m_indpxui_xw)
m_indrespt$lnincome<-haven::zap_labels(m_indrespt$lnincome)
m_indrespt$fasesXage<-haven::zap_labels(m_indrespt$fasesXage)
m_indrespt$fases<-haven::zap_labels(m_indrespt$fases)
m_indrespt$mases<-haven::zap_labels(m_indrespt$mases)

#as factor
m_indrespt$m_mastat_dv<-haven::as_factor(m_indrespt$m_mastat_dv)
m_indrespt$m_strata<-haven::as_factor(m_indrespt$m_strata)
m_indrespt$m_psu<-haven::as_factor(m_indrespt$m_psu)
m_indrespt$m_sex_dv<-haven::as_factor(m_indrespt$m_sex_dv)
m_indrespt$m_nchild_dv<-haven::as_factor(m_indrespt$m_nchild_dv)
m_indrespt$m_urban_dv<-haven::as_factor(m_indrespt$m_urban_dv)
m_indrespt$lvpab<-haven::as_factor(m_indrespt$lvpab)
m_indrespt$racel<-haven::as_factor(m_indrespt$racel)
m_indrespt$ledes<-haven::as_factor(m_indrespt$ledes)
m_indrespt$mawork <-haven::as_factor(m_indrespt$mawork)
m_indrespt$edu<-haven::as_factor(m_indrespt$edu)
m_indrespt$age_group<-haven::as_factor(m_indrespt$age_group)
m_indrespt$paedu <-haven::as_factor(m_indrespt$paedu )
m_indrespt$maedu  <-haven::as_factor(m_indrespt$maedu )
```

```{r}
#m_indrespimp<-sapply(m_indresptf, haven::zap_labels)

#factorial variables will considerably reduce the speed of imputation

#Wave 1 Number of logged events: 396; Time difference of 1.090519 hours
#Time difference of 5.846147 mins  Warning: Number of logged events: 250 
varnames_a<-c("a_mastat_dv","a_sex_dv","a_nchild_dv","a_urban_dv","mases","lvpab","edu","age_group","lnincome","racel","fases","fasesXage","paedu","maedu","mawork","leave","ledes")
start.time <- Sys.time() 

imputations_a<-mice(a_indrespt,pred=quickpred(a_indrespt,include= 
varnames_a,exclude= c("a_indinus_xw","a_strata","a_psu")))

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


#Wave 13 Number of logged events: 314 Time difference of 1.090526 hours
#Time difference of 5.846147 mins  Warning: Number of logged events: 250 
varnames_m<-c("m_mastat_dv","m_sex_dv","m_nchild_dv","m_urban_dv","mases","lvpab","edu","age_group","lnincome","racel","fases","fasesXage","paedu","maedu","mawork","leave","ledes")

start.time <- Sys.time() 

imputations_b<-mice(m_indrespt,pred=quickpred(m_indrespt,include= 
varnames_m,exclude= c("m_indpxui_xw","m_strata","m_psu")))

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken
```

```         
start.time <- Sys.time() ...Relevent codes... end.time <- Sys.time() time.taken <- end.time - start.time time.taken
```

### Wave 1 females 

```{r}
#with imputation: function built from https://gist.github.com/AaronGullickson/3ccb3fdd1778b32fc46df40d78faf5d3

lm_svy_af <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <- subset(svydesign(id=~a_psu, strata=~a_strata, 
                         weights=~a_indinus_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& a_sex_dv=="Female" )
    #run the model
    model <- svyglm(formula, design=imputation.svy,family=quasibinomial)
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, summary(model)$coef[,2])
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  #we can also grap n and p from the last model since 
  #they should be the same across all iterations
  n <- nobs(model)
  p <- length(model$coefficients)-1
  
  #return everything in a list
  return(list(coef=coefficients,
              n=n))
}


# for model 2 and 3 =============================================================
lm_svy_af_2 <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <- subset(svydesign(id=~a_psu, strata=~a_strata, 
                         weights=~a_indinus_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& a_sex_dv=="Female"  )
    #run the model
    model <- svy_vglm(formula, design=imputation.svy,family=multinomial(refLevel=1))
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, SE(model))
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  
  
  #return everything in a list
  return(list(coef=coefficients))
}


```

```{r}
#running models

#m1 Time difference of 5.143367 mins
start.time <- Sys.time() 

m1af_imp<-lm_svy_af(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,imputations_a)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken

#m2 Time difference of 18.08427 mins
start.time <- Sys.time() 

m2af_imp<-lm_svy_af_2(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,imputations_a)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


#m3 Time difference of 10.00591 mins
start.time <- Sys.time() 

m3af_imp<-lm_svy_af_2(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,imputations_a)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken

```

### Wave 1 males 

```{r}
#with imputation: function built from https://gist.github.com/AaronGullickson/3ccb3fdd1778b32fc46df40d78faf5d3

lm_svy_am <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <- subset(svydesign(id=~a_psu, strata=~a_strata, 
                         weights=~a_indinus_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& a_sex_dv=="Male"  )
    #run the model
    model <- svyglm(formula, design=imputation.svy,family=quasibinomial)
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, summary(model)$coef[,2])
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  #we can also grap n and p from the last model since 
  #they should be the same across all iterations
  n <- nobs(model)
  p <- length(model$coefficients)-1
  
  #return everything in a list
  return(list(coef=coefficients,
              n=n))
}


# for model 2 and 3 =============================================================
lm_svy_am_2 <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <- subset(svydesign(id=~a_psu, strata=~a_strata, 
                         weights=~a_indinus_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& a_sex_dv=="Male" )
    #run the model
    model <- svy_vglm(formula, design=imputation.svy,family=multinomial(refLevel=1))
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, SE(model))
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  
  
  #return everything in a list
  return(list(coef=coefficients))
}
```

```{r}
#m1 Time difference of 5.283394 mins
start.time <- Sys.time() 

m1am_imp<-lm_svy_am(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,imputations_a)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken

#m2 Time difference of 17.32093 mins
start.time <- Sys.time() 

m2am_imp<-lm_svy_am_2(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,imputations_a)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


#m3 Time difference of 10.53404 mins
start.time <- Sys.time() 

m3am_imp<-lm_svy_am_2(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,imputations_a)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken

```

### Wave 13 females

```{r}
#with imputation: function built from https://gist.github.com/AaronGullickson/3ccb3fdd1778b32fc46df40d78faf5d3

lm_svy_mf <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <- subset(svydesign(id=~m_psu, strata=~m_strata, 
                         weights=~m_indpxui_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& m_sex_dv=="Female" )
    #run the model
    model <- svyglm(formula, design=imputation.svy,family=quasibinomial)
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, summary(model)$coef[,2])
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  #we can also grap n and p from the last model since 
  #they should be the same across all iterations
  n <- nobs(model)
  p <- length(model$coefficients)-1
  
  #return everything in a list
  return(list(coef=coefficients,
              n=n))
}


# for model 2 and 3 =============================================================
lm_svy_mf_2 <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <- subset(svydesign(id=~m_psu, strata=~m_strata, 
                         weights=~m_indpxui_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& m_sex_dv=="Female" )
    #run the model
    model <- svy_vglm(formula, design=imputation.svy,family=multinomial(refLevel=1))
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, SE(model))
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  
  
  #return everything in a list
  return(list(coef=coefficients))
}


```

```{r}
#running models

#m1 Time difference of 2.939297 mins
start.time <- Sys.time() 

m1bf_imp<-lm_svy_mf(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,imputations)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken

#m2 Time difference of 8.87294 mins
start.time <- Sys.time() 

m2bf_imp<-lm_svy_mf_2(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,imputations)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


#m3 Time difference of 4.866161 mins
start.time <- Sys.time() 

m3bf_imp<-lm_svy_mf_2(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,imputations)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


```

### Wave 13 males

```{r}
#with imputation: function built from https://gist.github.com/AaronGullickson/3ccb3fdd1778b32fc46df40d78faf5d3

lm_svy_mm <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <-subset(svydesign(id=~m_psu, strata=~m_strata, 
                         weights=~m_indpxui_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& m_sex_dv=="Male")
    
    #run the model
    model <- svyglm(formula, design=imputation.svy,family=quasibinomial)
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, summary(model)$coef[,2])
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  #we can also grap n and p from the last model since 
  #they should be the same across all iterations
  n <- nobs(model)
  p <- length(model$coefficients)-1
  
  #return everything in a list
  return(list(coef=coefficients,
              n=n))
}


# for model 2 and 3 =============================================================
lm_svy_mm_2 <- function(formula, imputations) {
  options(survey.lonely.psu="adjust")
  
  #setting up null objects allows us to easily add results
  #later
  b <- se<- NULL
  
  #now loop through our imputations and run the model
  for(i in 1:imputations$m) {
    #grab the complete dataset
    imputation <- complete(imputations, i)
    #create the design effect object
    imputation.svy <-subset(svydesign(id=~m_psu, strata=~m_strata, 
                         weights=~m_indpxui_xw, data=imputation),age_group%in%c("16-20","21-25","26-30","31-35")& m_sex_dv=="Male")
        
    #run the model
    model <- svy_vglm(formula, design=imputation.svy,family=multinomial(refLevel=1))
    #collect the results
    b <- cbind(b, coef(model))
    se <- cbind(se, SE(model))
  }
  
  #now pool the results
  b.pool <- apply(b, 1, mean)
  between.var <- apply(b, 1, var)
  within.var <- apply(se^2, 1, mean)
  se.pool <- sqrt(within.var+between.var+between.var/imputations$m) 
  t.pool <- b.pool/se.pool 
  pvalue.pool <- (1-pnorm(abs(t.pool)))*2 
  coefficients <- data.frame(b.pool, se.pool, t.pool, pvalue.pool)
  
  
  #return everything in a list
  return(list(coef=coefficients))
}



```

```{r}
#m1 Time difference of 3.179687 mins
start.time <- Sys.time() 

m1bm_imp<-lm_svy_mm(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,imputations)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


#m2 Time difference of 8.01764 mins
start.time <- Sys.time() 

m2bm_imp<-lm_svy_mm_2(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,imputations)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


#m3 Time difference of 4.359161 mins
start.time <- Sys.time() 

m3bm_imp<-lm_svy_mm_2(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,imputations)

end.time <- Sys.time() 
time.taken <- end.time - start.time 
time.taken


```

```{r}
fitm1b <- with(imputations, glm(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,family=binomial(link="logit")))
summary(pool(fitm1b))
```

## Plotting results

```{r}
convertModel <- function(model) {
  tr <- createTexreg(
    coef.names = rownames(model$coef), 
    coef = model$coef$b.pool, 
    se = model$coef$se.pool, 
    pvalues = model$coef$pvalue.pool,
  )
}
#compare males and females in wave 1
 af_m1<-screenreg(lapply(list(m1af_imp, m1am_imp), convertModel))
 af_m2<-screenreg(lapply(list(m2af_imp, m2am_imp), convertModel))
 af_m3<-screenreg(lapply(list(m3af_imp, m3am_imp), convertModel),file="compare males and females in wave 1.doc",model.names = c("Wave 1", "Wave 13"),omit.coef =c("(Intercept):1", "(Intercept):2", "age_group21-25:1","age_group21-25:2","age_group26-30:1","age_group26-30:2" ,"age_group31-35:1","age_group31-35:2","factor(edu)middle:1", "factor(edu)middle:2", "factor(edu)high:1", "factor(edu)high:2", "factor(racel)African Caribbean:1",   "factor(racel)African Caribbean:2" , "factor(racel)Indiani,Pakistani, Bangladeshi:1", "factor(racel)Indiani,Pakistani, Bangladeshi:2" ,   "factor(racel)Other Asian:1",     "factor(racel)Other Asian:2",
"factor(racel)Other:1", "factor(racel)Other:2", "m_urban_dvrural area:1", "m_urban_dvrural area:2" ))

#compare males and females in wave 13
 mf_m1<-screenreg(lapply(list(m1bf_imp, m1bm_imp), convertModel))
 mf_m2<-screenreg(lapply(list(m2bf_imp, m2bm_imp), convertModel))
 mf_m3<-screenreg(lapply(list(m3bf_imp, m3bm_imp), convertModel))
 
#compare females in wave 13 and wave 1
 bf_m1<-screenreg(lapply(list(m1af_imp, m1bf_imp), convertModel))
 bf_m2<-screenreg(lapply(list(m2af_imp, m2bf_imp), convertModel))
 bf_m3<-screenreg(lapply(list(m3af_imp, m3bf_imp), convertModel))
 
 #compare males in wave 13 and wave 1
 bm_m1<-screenreg(lapply(list(m1am_imp, m1bm_imp), convertModel))
 bm_m2<-screenreg(lapply(list(m2am_imp, m2bm_imp), convertModel))
 bm_m3<-screenreg(lapply(list(m3am_imp, m3bm_imp), convertModel))
        
```

## Assumption checking

### imputation

 <https://stats.oarc.ucla.edu/r/faq/how-do-i-perform-multiple-imputation-using-predictive-mean-matching-in-r/>

```{r}
a_indrespy<-a_indresp[which(a_indresp$a_age_dv%in%c(16:35)),]
m_indrespy<-m_indresp[which(m_indresp$m_age_dv%in%c(16:35)),]
```

```{r}
marginplot(a_indrespy[c(17, 20)], col = c("blue", "red", "orange"))
```

The box plots are blue for observed values and red for missing values

Evidence supporting MAR over MCAR would be if the distribution of **x2** for those observations with missing information for **y1** or **y4** were much higher or much lower than those of the non-missing observations. In the above graph, the boxplots appear to mostly overlap once again providing support for the assumption of MCAR.

```{r}
pbox(a_indrespy, pos = 17)
```

```{r}
pbox(m_indresp, pos = 17)
```

# Model building without imputations

## Model 1

```{r}
#for females
#without imputation 
#females
m1af<-svyglm(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,design=svy_indresp1yf,na.action=na.omit,family=quasibinomial)

m1bf<-svyglm(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,design=svy_indresp2yf,na.action=na.omit,family=quasibinomial)

#males
m1am<-svyglm(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,design=svy_indresp1ym,na.action=na.omit,family=quasibinomial)

m1bm<-svyglm(lvpab~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,design=svy_indresp2ym,na.action=na.omit,family=quasibinomial)

```

```{r}
#z tests of coefficients
zscore <- function(coef2,coef1,se2,se1){(coef2-coef1)/sqrt((se2)**2+(se1)**2)
}
#females 
#fases
zscore(coef2 =    1.35783  , coef1=  0.493502       , se2= 0.46276 , se1= 0.084185 )
#maedu (high)
zscore(coef2 = -0.16265      ,coef1= -0.484515      , se2=0.53702   , se1=0.202592)
#paedu (high)
zscore(coef2 =   0.50519       ,coef1=  -0.200648       , se2=0.57399   , se1=0.185489  )
#mawork sig
zscore(coef2 =  -0.38087             , coef1=  0.359646         , se2= 0.31075    , se1= 0.113000  )


#males 
#fases
zscore(coef2 =  0.77227           ,coef1=   0.292779         , se2=  0.66703 , se1=  0.093648   )
#maedu (high)
zscore(coef2 =    -0.79371           ,coef1=-0.188297        , se2=0.63461   , se1=0.218553)
#paedu (high)
zscore(coef2 =   -0.19252             ,coef1= -0.320714          , se2=0.58333   , se1=0.218249    )
#mawork sig
zscore(coef2 =  -0.63024              , coef1= 0.181253           , se2= 0.35183    , se1=  0.131726  )


```

```{r}
#ames
#females difference -0.122  p.value  0.021
cames<-summary(marginaleffects(m1af,variables="mawork"))
cames2<-summary(marginaleffects(m1bf,variables="mawork"))
came<-rbind(cames,cames2)
compare.margins(came$estimate,came$std.error)

#males difference -0.141; p.value  0.017
cames<-summary(marginaleffects(m1am,variables="mawork"))
cames2<-summary(marginaleffects(m1bm,variables="mawork"))
came<-rbind(cames,cames2)
compare.margins(came$estimate,came$std.error)
```

## Model 2

```{r}
m2af<-svy_vglm(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,family=multinomial(refLevel=1),design=svy_indresp1yf,na.action=na.omit)

m2bf<-svy_vglm(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,family=multinomial(refLevel=1),design=svy_indresp2yf,na.action=na.omit)

m2am<-svy_vglm(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,family=multinomial(refLevel=1),design=svy_indresp1ym,na.action=na.omit)

m2bm<-svy_vglm(ledes~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,family=multinomial(refLevel=1),design=svy_indresp2ym,na.action=na.omit)

```

```{r}
#z test for coefficients
#females
##cohabitation
###pases
zscore(coef2 =  -1.145945               ,coef1=   -0.5279231          , se2=  0.578323 , se1= 0.1067982    )
###maedu(high)
zscore(coef2 =  0.535709             ,coef1=    0.3910630          , se2=  0.647512   , se1=  0.2386345   )
###paedu(high)
zscore(coef2 = -0.408988           ,coef1=   -0.0225736         , se2=  0.690724, se1= 0.2202824   )
###mawork 
zscore(coef2 =  0.247904          ,coef1=   -0.2549268        , se2=  0.359905 , se1=  0.1396798   )

##marriage
###pases
zscore(coef2 = -1.358326          ,coef1=   -0.7566254          , se2=  0.563529 , se1= 0.1290402   )
###maedu(high)
zscore(coef2 =   0.358469            ,coef1=    0.2141034         , se2=  0.612275 , se1=  0.2432039   )
###paedu(high)
zscore(coef2 =   -0.120479          ,coef1=    0.3031643          , se2=  0.632659  , se1=  0.2209672     )
###mawork sig=1.97 not sig coefficient
zscore(coef2 = 0.391774           ,coef1=  -0.3863171          , se2=  0.371241 , se1=  0.1345737   )

##without a partner
###pases
zscore(coef2 =  -1.604960        ,coef1=    -0.4148343           , se2=  0.626037 , se1= 0.0975044    )
###maedu(high)
zscore(coef2 =    -0.165066           ,coef1=    0.6751570        , se2=   0.606883 , se1=  0.2180539   )
###paedu(high)
zscore(coef2 =    -0.915336          ,coef1=     0.2940644          , se2=   0.634411  , se1=   0.2040251     )
###mawork sig=1.97 sig cof for wave 1, not sig for wave 13
zscore(coef2 =  0.384831           ,coef1=  -0.4305743          , se2=  0.394867 , se1=  0.1235926   )


#males
##cohabitation
###pases
zscore(coef2 = -0.1573351         ,coef1=    -0.4207956           , se2=  0.8356579  , se1= 0.1462146     )
###maedu(high)
zscore(coef2 =    0.7581093              ,coef1=    0.0648927       , se2=   0.8468744 , se1=  0.2569212   )
###paedu(high)
zscore(coef2 =    -0.0895241           ,coef1=      -0.1840590         , se2=    0.7805476   , se1=   0.2627747     )
###mawork sig=2.09847 not sig coefficient
zscore(coef2 =   1.0827787             ,coef1=  -0.1310800           , se2=   0.5570283  , se1=   0.1559587   )

##marriage
###pases
zscore(coef2 = -0.5434834        ,coef1=     -1.0634425           , se2=  0.8331357  , se1= 0.1632331     )
###maedu(high)
zscore(coef2 =     1.4201381               ,coef1=   0.2325718        , se2=    0.7789965   , se1=  0.2625822   )
###paedu(high)
zscore(coef2 =    -0.1181511           ,coef1=      0.1469352         , se2=    0.7352541   , se1=   0.2439483     )
###mawork
zscore(coef2 =    0.3984417           ,coef1=  -0.0598869            , se2=  0.4306555   , se1=   0.1543132   )


##without a partner
###pases
zscore(coef2 = -1.1276304         ,coef1=     -0.1783426          , se2=  0.7328520 , se1= 0.1021658      )
###maedu(high)
zscore(coef2 =      0.4877030                ,coef1=    0.2378884         , se2=     0.7266739   , se1=   0.2431004       )
###paedu(high)
zscore(coef2 =     0.6922228           ,coef1=       0.6772597          , se2=    0.6848506    , se1=   0.2416268     )
###mawork
zscore(coef2 =    0.5349542             ,coef1= -0.3293337            , se2=  0.4310773   , se1=    0.1505320    )

```

```{r}
#m2af sig 
#fases:1,2,3 -
#factor(mawork)working:1,3 -
#factor(maedu)high:3 +


#m2bf sig
#fases:1,2,3 -

#m2am sig
#fases:1,2,3 -
#factor(mawork)working:3 -

#m2bm sig none

```

## Model 3

```{r}
m3af<-svy_vglm(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,family=multinomial(refLevel=1),design=svy_indresp1yf,na.action=na.omit)

m3bf<-svy_vglm(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,family=multinomial(refLevel=1),design=svy_indresp2yf,na.action=na.omit)

m3am<-svy_vglm(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+a_urban_dv,family=multinomial(refLevel=1),design=svy_indresp1ym,na.action=na.omit)

m3bm<-svy_vglm(leave~fases+fasesXage+age_group+factor(mawork)+factor(maedu)+factor(paedu)+factor(edu)+factor(racel)+m_urban_dv,family=multinomial(refLevel=1),design=svy_indresp2ym,na.action=na.omit)
```

```{r}
#z test for coefficients
#females
##with partners
###pases
zscore(coef2 =     -1.4293392                  ,coef1=  -0.4570019             , se2=  0.5268451    , se1=   0.0878269   )
###maedu(high)

###paedu(high)

###mawork 


```

```{r}
#m3bm sig
#mawork 2 +

#m3am sig
#fases 1,2 -


#m3bf
#fases 1,2 -

#m3af
#fases 1,2-
#mawork 1,2-
#maedu high +
```

```{r}
#whether mother working when male respondents age 14 have higher predicability of home-leaving without a partner in wave 13 than wave 1
#mawork sig at wave 13 not wave 1
zscore(coef2=   0.799274      ,coef1= -0.2217087        ,se2= 0.395233   ,se1= 0.1357911)

#for females the effect direction is reversed, with effects only sig at wave 1
zscore(coef2=    0.3330332      ,coef1=  -0.3454488         ,se2= 0.3347052   ,se1=  0.1172492)

```

```{r}

testa<-a_indresptf%>%dplyr::select(leave,fases,fasesXage,age_group,mawork,maedu,paedu,edu,racel,a_urban_dv,a_sex_dv)
test1<-complete.cases(testa)
table(test1)

testb<-m_indrespt%>%dplyr::select(leave,fases,fasesXage,age_group,mawork,maedu,paedu,edu,racel,m_urban_dv,m_sex_dv)
test2<-complete.cases(testb)
table(test2)

#comple cases of males and females in wave 1 and 13: 434, 3927
```
