---
title: "UKHLS exploration (kins with weights) ukhls13"
format: html
editor: visual
---

```{r,cache=T}
##reading ukhls data with masee variables
setwd("D:\\r git projects\\ox-R\\final essay\\Understanding Society Waves 2009-2022 and Harmonised BHPS Waves 1991-2009\\UKDA-6614-tab\\tab\\ukhls")
#ukhls_indi1<-read.table("a_indresp.tab",header=T)

ukhls_indi13<-read.table("m_indresp.tab",header=T)
```

# Cleaning Wave 13

```{r}
#weighting the ukhls13 with m_indpxui_xw
library(tidyverse)
library(survey)
library(svrep)
library(Hmisc)
library(srvyr)
library(marginaleffects)
library(haven)
library(catregs)
library(margins)
library(modelsummary)

#no missing in lvrel
ukhls_indi13$m_lvrel1[ukhls_indi13$m_lvrel1%in%c(-7,-2,-1)]<-NA
ukhls_indi13$m_lvrel2[ukhls_indi13$m_lvrel2%in%c(-7,-2,-1)]<-NA
ukhls_indi13$m_lvrel9[ukhls_indi13$m_lvrel9%in%c(-7,-2,-1)]<-NA
ukhls_indi13$m_lvrel10[ukhls_indi13$m_lvrel10%in%c(-7,-2,-1)]<-NA

#assigning pns values
ukhls_indi13$m_pns1pno[ukhls_indi13$m_pns1pno!=0]<-1
ukhls_indi13$m_pns2pno[ukhls_indi13$m_pns2pno!=0]<-1

#num of parents (not censored)
ukhls_indi13$parents<-ukhls_indi13$m_lvrel1+ukhls_indi13$m_lvrel2+ukhls_indi13$m_lvrel9+ukhls_indi13$m_lvrel10+ukhls_indi13$m_pns1pno+ukhls_indi13$m_pns2pno

#having one & two or more parents (censored)
ukhls_indi13$panum<-c(0)
ukhls_indi13$panum[ukhls_indi13$parents==1]<-1
ukhls_indi13$panum[ukhls_indi13$parents>1]<-2

#living with parents or not
ukhls_indi13$lvpa<-ukhls_indi13$m_pns1pno+ukhls_indi13$m_pns2pno
#binary
ukhls_indi13$lvpab<-c(0)
ukhls_indi13$lvpab[ukhls_indi13$lvpa!=0]<-1


#no missing in age
ukhls_indi13$m_age_dv[ukhls_indi13$m_age_dv==-9]<-NA

#no missing in sex
ukhls_indi13$m_sex_dv[ukhls_indi13$m_sex_dv%in%c(-9,0)]<-NA

#no missing in education
ukhls_indi13$m_qfhigh_dv[ukhls_indi13$m_qfhigh_dv%in%c(-9,-8,0)]<-NA

#high ses and low ses
ukhls_indi13$ses<-c(0)
ukhls_indi13$ses[ukhls_indi13$m_qfhigh_dv%in%c(1,2,3,4,5,6)]<-1


#no missing in nchild_dv and nchresp
#table(ukhls_indi13$m_nchild_dv)
ukhls_indi13$m_nchresp[ukhls_indi13$m_nchresp%in%c(-7)]<-NA

#make a variable of Children under 16 not living in HH
ukhls_indi13$ohch16<-c(0)
ukhls_indi13$ohch16[ukhls_indi13$m_ohch16%in%c(1,2)]<-1


#making nchild_dv a binary variable
ukhls_indi13$child<-c(0)
ukhls_indi13$child[ukhls_indi13$m_nchild_dv>=1]<-1

#no missing in mastat_dv
ukhls_indi13$m_mastat_dv[ukhls_indi13$m_mastat_dv%in%c(-9,-2,-1)]<-NA

#making partner a binary variable
ukhls_indi13$partner<-c(0)
ukhls_indi13$partner[ukhls_indi13$m_mastat_dv%in%c(2,3,10)]<-1

#makeing cohab a binary variable
ukhls_indi13$cohab<-c(0)
ukhls_indi13$cohab[ukhls_indi13$m_mastat_dv%in%c(10)]<-1

#making married a binary variable
ukhls_indi13$mar<-c(0)
ukhls_indi13$mar[ukhls_indi13$m_mastat_dv%in%c(2,3)]<-1
#child-bearing within wedlock
ukhls_indi13$wedch<-c(0)
ukhls_indi13$wedch[ukhls_indi13$mar==1 & ukhls_indi13$m_nchild_dv !=0]<-1


#urban
#table(ukhls_indi13$m_urban_dv)
ukhls_indi13$m_urban_dv[ukhls_indi13$m_urban_dv%in%c(-9)]<-NA


#lnage
ukhls_indi13$lnage<-log(ukhls_indi13$m_age_dv)

#income
ukhls_indi13$m_fimngrs_dv[ukhls_indi13$m_fimngrs_dv<0]<-NA
#lnincome
ukhls_indi13$lnincome<-log(ukhls_indi13$m_fimngrs_dv)
ukhls_indi13$lnincome[ukhls_indi13$lnincome==-Inf]<-NA


#unmarried women
#married women
```

# Route out (multinomial):

nchresp responsible for a child

nchild_dv have a child in the hh

whether son/daughter alive: using lvrel 3!!!

-   **ohch 16**: children under 16 not living in hh

Are any of your children living outside the household aged under 16?

#### **Question universe:**

if LVRel = 3 (If son(s)/daughter(s) alive)

**consider whether father living with their children under 16 (n=220), but not consider mothers (n=49) within the age range 25 to 50**

```{r}
##patterns of home-leaving 
ukhls_indi13$homele<-c(1)

##route out 1: to live with a partner and children ========================
###f
ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                            ukhls_indi13$m_sex_dv==2 & #females
                            ukhls_indi13$m_mastat_dv%in%c(2,3,10)& #married or cohabiting
                            #ukhls_indi13$ohch16==0 & # no child under 16 not living with res
                            ukhls_indi13$m_nchild_dv!=0]<-2 #have a child under 16 in the hh
###m
ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                            ukhls_indi13$m_sex_dv==1 & #males
                            ukhls_indi13$m_mastat_dv%in%c(2,3,10)& #married or cohabiting
                            ukhls_indi13$ohch16==0 & # no child under 16 not living with res
                            ukhls_indi13$m_nchild_dv!=0]<-2 #have a child under 16 in the hh

##route out 2: to live with a partner but no children=====================
###f
ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                            ukhls_indi13$m_sex_dv==2 & #females
                            ukhls_indi13$m_mastat_dv%in%c(2,3,10) & #married or cohabiting
                            #ukhls_indi13$ohch16==0 & # no child under 16 not living with res
                            ukhls_indi13$m_nchild_dv==0]<-3 # don't have a child under 16 in the hh

###m
ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                            ukhls_indi13$m_sex_dv==1 & #males
                            ukhls_indi13$m_mastat_dv%in%c(2,3,10) & #married or cohabiting
                            ukhls_indi13$ohch16==0 & # no child under 16 not living with res
                            ukhls_indi13$m_nchild_dv==0]<-3 # don't have a child under 16 in the hh


##route out 3: forming a female-headed family============================
##for females
ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                    ukhls_indi13$m_sex_dv==2 & #females
                            ukhls_indi13$m_mastat_dv!=2 & #not married
                            ukhls_indi13$m_mastat_dv!=3 & #not married
                            ukhls_indi13$m_mastat_dv!=10 & #no partner
                            ukhls_indi13$m_nchild_dv!=0]<-4 #have a dep child in hh


##for males
ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                    ukhls_indi13$m_sex_dv==1 & #males
                    ukhls_indi13$ohch16==1 # (at least one) children under 16 not living in hh
                    ]<-4


##route out 4: to autonomous living =====================
###f
ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                            ukhls_indi13$m_sex_dv==2 & #females
                            ukhls_indi13$m_mastat_dv!=2 & #not married
                            ukhls_indi13$m_mastat_dv!=3 & #not married
                            ukhls_indi13$m_mastat_dv!=10 & #no partner
                            #ukhls_indi13$ohch16==0 & # no child under 16 not living with res
                            ukhls_indi13$m_nchild_dv==0]<-5 # not have a dep in the hh

ukhls_indi13$homele[ukhls_indi13$lvpa==0 & # not living with parents
                            ukhls_indi13$m_sex_dv==1 & #males
                            ukhls_indi13$m_mastat_dv!=2 & #not married
                            ukhls_indi13$m_mastat_dv!=3 & #not married
                            ukhls_indi13$m_mastat_dv!=10 & #no partner
                            ukhls_indi13$ohch16==0 & # no child under 16 not living with res
                            ukhls_indi13$m_nchild_dv==0]<-5 # not have a dep in the hh


# check:table(ukhls_indi13$m_nchild_dv,ukhls_indi13$m_nchresp)

```

```{r}
#for females who have children under 16 not living with them
test<-ukhls_indi13[which(ukhls_indi13$lvpa==0 & ukhls_indi13$m_sex_dv==2 & ukhls_indi13$ohch16==1),]
test2<-ukhls_indi13[which(ukhls_indi13$lvpa==0 & ukhls_indi13$m_sex_dv==1 & ukhls_indi13$ohch16==1),]

dim(test)
dim(test2)

table(ukhls_indi13$m_sex_dv,ukhls_indi13$ohch16)
 
##=======================================================for single parents males n females
sin<-ukhls_indi13[which(ukhls_indi13$lvpa==0 & ukhls_indi13$m_sex_dv==2& ukhls_indi13$m_mastat_dv!=2& ukhls_indi13$m_mastat_dv!=3& ukhls_indi13$m_mastat_dv!=10 & ukhls_indi13$m_nchild_dv!=0),]

sin2<-ukhls_indi13[which(ukhls_indi13$lvpa==0 & ukhls_indi13$m_sex_dv==1& ukhls_indi13$m_mastat_dv!=2& ukhls_indi13$m_mastat_dv!=3& ukhls_indi13$m_mastat_dv!=10 & ukhls_indi13$m_nchild_dv!=0),]

##===============================still use female-headed family if father live separately with the child
age<-ukhls_indi13[which(ukhls_indi13$m_age_dv%in%c(25:50)),]
table(age$ohch16,age$m_sex_dv)
table(age$homele,age$m_sex_dv)

##most of the depedent children not living with females are their grandchildren!!!
table(test$m_age_dv)

##for females who not living with their children under 16 age 25 to 50 (num 49, compared with men 220)
agefab<-age[which(age$m_sex_dv==2 & age$ohch16==1),]
table(agefab$m_mastat_dv)
table(agefab$lvpa)

##check whether all fathers not living with their children are included
c<-age[which(age$lvpa==0 & age$m_sex_dv==1 & age$ohch16==1),]
```

# Nonmarital childbearing (multinomial)

no children

have children under 16 (no married)

have children under 16 (married)

```{r}

```

```{r}
##adopting the complex survey design
##??
options(survey.lonely.psu = "adjust")
design<-svydesign(id=~m_psu, weights=~m_indinui_xw,strata =~m_strata,data=ukhls_indi13)
#design<-ukhls_indi13%>%srvyr::as_survey(weights=m_indpxui_xw)


#subset the data for respondents aged 25-50
ukhls13<-subset(design, m_age_dv%in%c(25:50))
#ukhls13_child<-subset(ukhls13, child%in%c(1))


```

## Model building

### multinomial logit

```{r}
library(svyVGAM)
#for home-leaving patterns
m0<-svyVGAM::svy_vglm(homele~ses*m_sex_dv+ses*panum+lnage+m_urban_dv+lnincome,family=multinomial(refLevel=1), design=ukhls13,na.action=na.omit) 

```

Drafts

```{r}
m<-svyglm(parents~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit)
m2<-svyglm(m_nchild_dv~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit,family=poisson())

m3<-svyglm(partner~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit)

m4<-svyglm(child~ses+m_age_dv+m_sex_dv,family=quasibinomial,design=ukhls13,na.action = na.omit)

#m4<-svyglm(panum~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit)

#logistic regression for binary depedent variable
m5<-svyglm(partner~ses+m_age_dv+m_sex_dv,family=quasibinomial,design=ukhls13,na.action = na.omit)

#possion regression for count model
m6<-svyglm(parents~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit,family=poisson())
m66<-svyglm(parents~ses+m_age_dv+m_sex_dv,design=ukhls13_child,na.action = na.omit,family=poisson())

m7<-svyglm(panum~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit,family=poisson())

m8<-svyglm(lvpa~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit,family=poisson())

m9<-svyglm(lvpab~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit,family=quasibinomial)
m99<-svyglm(lvpab~ses*m_sex_dv+m_age_dv,design=ukhls13,na.action = na.omit,family=quasibinomial)

#for parent's divorce? no meaningful results
#m10<-svyglm(lvpab~ses+m_age_dv+m_sex_dv+factor(lvpa),design=ukhls13,na.action = na.omit,family=quasibinomial)

#for cohab for high n low ses
m10<-svyglm(cohab~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit,family=quasibinomial)

#for whether child-bearing within wedlock
m11<-svyglm(wedch~ses+m_age_dv+m_sex_dv,design=ukhls13_child,na.action = na.omit,family=quasibinomial) 

#m111<-svyglm(wedch~ses+m_age_dv+m_sex_dv,design=ukhls13,na.action = na.omit,family=quasibinomial)

#for out-of and within wedlock child-bearing
m12<-svyglm(child~ses+m_age_dv+m_sex_dv+mar,design=ukhls13,na.action = na.omit,family=quasibinomial)
```

```{r}
##plotting for partner
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(m5,newdata=dm,type="response",se=T))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```

```{r}
##plotting for child
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(m4,newdata=dm,type="response",se=T))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```

```{r}
testukhls<-ukhls_indi13[which(ukhls_indi13$m_age_dv%in%c(25:50)),]

test<-glm(child~ses+m_age_dv+m_sex_dv,data=testukhls,family= binomial,na.action = na.omit)

##plotting for child (not weighted)
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(test,newdata=dm,type="response"))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```

```{r}
##plotting for panum
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(m7,newdata=dm,type="response",se=T))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```

```{r}
##plotting for parents
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(m6,newdata=dm,type="response",se=T))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```

```{r}
##plotting for parents
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(m2,newdata=dm,type="response",se=T))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```

```{r}
##plotting for lvpa
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(m8,newdata=dm,type="response",se=T))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```

```{r}
##plotting for lvpab
dm<-data.frame(m_sex_dv=rep(c(1,2),each=26),ses=rep(c(1,0),26),m_age_dv=rep(c(25:50),2))

library(reshape2)
pdm<-cbind(dm,predict(m9,newdata=dm,type="response",se=T))

ldpm<-melt(pdm,id.vars=c("m_sex_dv","ses","m_age_dv"),value.name="probability")

ldpm_F<-ldpm[which(ldpm$m_sex_dv==2),]
ldpm_M<-ldpm[which(ldpm$m_sex_dv==1),]

ggplot(ldpm_F,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()

ggplot(ldpm_M,aes(x=m_age_dv,y=probability,color=variable,linetype=as.factor(ses)))+geom_line()+theme_minimal()
```
