---
title: "UKHLS exploration (weighed kin) 4.20"
format: html
editor: visual
---

## Reading and Cleaning data

```{r echo = T, results = 'hide'}
#houskeeping
#Clear objects already in the environment – start with a clean slate
rm(list=ls())

#loading libraries
library(tidyverse)
library(svyVGAM)
library(sjlabelled)
library(desctable)
library(summarytools)
library(naniar)
library(survey)
library(svrep)
library(Hmisc)
library(srvyr)
library(marginaleffects)
library(haven)
library(catregs)
library(margins)
library(modelsummary)
library(zoo)
library(mice)
library(stargazer)
library(texreg)
library(VIM)
library(lattice)
library(ggplot2)
```

### Wave 1-5 egoalt (8-6) hhresp (8-6)

```{r}
inpath<-"D:/r git projects/ox-R/final essay/UKHLS n BHPS stata/UKDA-6614-stata/stata/stata13_se/"
missval <- c(-9, -8, -7, -2, -1)

```

### Wave 8\~13 egoalt (9-13) hhresp (9-13)

```{r}
#data================================================
#wave 8
h_indresp <- read_dta(file=paste0(inpath, "ukhls/h_indresp.dta")) 
h_indresp<-h_indresp%>%dplyr::select(pidp, h_hidp, h_mastat_dv,
         h_age_dv, h_sex_dv, h_hiqual_dv,
         h_mastat_dv, h_racel_dv,h_fimnnet_dv,h_nchild_dv,h_urban_dv,h_gor_dv,h_jbstat,h_sf1,h_scsf1)

#wave 9
i_indresp <- read_dta(file=paste0(inpath, "ukhls/i_indresp.dta"))
i_indresp<-i_indresp%>%dplyr::select(pidp, i_hidp, i_mastat_dv,
         i_age_dv, i_sex_dv, i_hiqual_dv,
         i_mastat_dv, i_racel_dv,i_fimnnet_dv,i_nchild_dv,i_urban_dv,i_gor_dv,i_jbstat,i_sf1,i_scsf1)

#wave 10
j_indresp <- read_dta(file=paste0(inpath, "ukhls/j_indresp.dta"))
j_indresp<-j_indresp%>%dplyr::select(pidp, j_hidp, j_mastat_dv,
         j_age_dv, j_sex_dv, j_hiqual_dv,
         j_mastat_dv, j_racel_dv,j_fimnnet_dv,j_nchild_dv,j_urban_dv,j_gor_dv,j_jbstat,j_sf1,j_scsf1)


#wave 11
k_indresp <- read_dta(file=paste0(inpath, "ukhls/k_indresp.dta"))
k_indresp<-k_indresp%>%dplyr::select(pidp, k_hidp, k_mastat_dv,
         k_age_dv, k_sex_dv, k_hiqual_dv,
         k_mastat_dv, k_racel_dv,k_fimnnet_dv,k_nchild_dv,k_urban_dv,k_gor_dv,k_jbstat,k_sf1,k_scsf1)

#wave 12
l_indresp <- read_dta(file=paste0(inpath, "ukhls/l_indresp.dta"))
l_indresp<-l_indresp%>%dplyr::select(pidp, l_hidp, l_mastat_dv,
         l_age_dv, l_sex_dv, l_hiqual_dv,
         l_mastat_dv, l_racel_dv,l_fimnnet_dv,l_nchild_dv,l_urban_dv,l_gor_dv,l_jbstat,l_sf1,l_scsf1)

#wave 13
m_indresp <- read_dta(file=paste0(inpath, "ukhls/m_indresp.dta"))

m_indresp<-m_indresp%>%dplyr::select(pidp, m_hidp, m_mastat_dv,
         m_age_dv, m_sex_dv, m_hiqual_dv,
         m_mastat_dv, m_racel_dv,m_fimnnet_dv,m_nchild_dv,m_urban_dv,m_gor_dv,m_jbstat,m_sf1,m_scsf1)



#egoalt================================================
h_egoalt<-read_dta(file=paste0(inpath, "ukhls/h_egoalt.dta")) 

i_egoalt<-read_dta(file=paste0(inpath, "ukhls/i_egoalt.dta")) 

j_egoalt<-read_dta(file=paste0(inpath, "ukhls/j_egoalt.dta")) 

k_egoalt<-read_dta(file=paste0(inpath, "ukhls/k_egoalt.dta")) 

l_egoalt<-read_dta(file=paste0(inpath, "ukhls/l_egoalt.dta")) 

m_egoalt<-read_dta(file=paste0(inpath, "ukhls/m_egoalt.dta")) 

#hhresp================================================
h_hhresp<-read_dta(file=paste0(inpath, "ukhls/h_hhresp.dta"))  %>%
 dplyr::select(h_hidp,h_tenure_dv)

i_hhresp<-read_dta(file=paste0(inpath, "ukhls/i_hhresp.dta"))  %>%
 dplyr::select(i_hidp,i_tenure_dv)

j_hhresp<-read_dta(file=paste0(inpath, "ukhls/j_hhresp.dta"))  %>%
 dplyr::select(j_hidp,j_tenure_dv)

k_hhresp<-read_dta(file=paste0(inpath, "ukhls/k_hhresp.dta"))  %>%
 dplyr::select(k_hidp,k_tenure_dv)

l_hhresp<-read_dta(file=paste0(inpath, "ukhls/l_hhresp.dta"))  %>%
 dplyr::select(l_hidp,l_tenure_dv)

m_hhresp<-read_dta(file=paste0(inpath, "ukhls/m_hhresp.dta"))  %>%
 dplyr::select(m_hidp,m_tenure_dv)
```

```{r}
##remove all missing values in indresp
for (i in 1:5) {
  h_indresp<- h_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  i_indresp<- i_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  j_indresp<- j_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  k_indresp<- k_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))
}
for (i in 1:5) {
  l_indresp<- l_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))}

for (i in 1:5) {
  m_indresp<- m_indresp %>%
    mutate_all(., list(~na_if(., missval[i])))}
```

### Getting the youths not living with parents in the next wave

#### Wave 1\~5

```{r}

```

#### Wave 8\~12

##### Wave 8

```{r}
mothers <- h_egoalt %>%
  filter(h_relationship_dv %in% c(4) & h_asex==2) %>%
  select(pidp, apidp)
mothers$mother_pidp=mothers$apidp

h_egoalt <- h_egoalt %>%
  full_join(mothers, by=c("pidp", "apidp")) 

rm(mothers)

fathers <- h_egoalt %>%
  filter(h_relationship_dv %in% c(4) & h_asex==1) %>%
  select(pidp, apidp)
fathers$father_pidp=fathers$apidp

h_egoalt <- h_egoalt %>%
  full_join(fathers, by=c("pidp", "apidp"))

rm(fathers)


#wave 8 leave: not live with mothers
h_egoalt$separated_m <- labelled(
  ifelse((!is.na(h_egoalt$mother_pidp) & h_egoalt$h_enwstat==5), -8, 
  ifelse((!is.na(h_egoalt$mother_pidp)  & h_egoalt$h_anwstat==5), 3,
  ifelse((!is.na(h_egoalt$mother_pidp)  & 
          h_egoalt$h_anwstat==2 & h_egoalt$h_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(h_egoalt$mother_pidp) & 
          h_egoalt$h_anwstat %in% c(3,4) & h_egoalt$h_enwstat==2), 2,
  ifelse((!is.na(h_egoalt$mother_pidp) &
          h_egoalt$h_anwstat==2 & h_egoalt$h_enwstat==2), 1,
  ifelse((!is.na(h_egoalt$mother_pidp)& h_egoalt$h_anwstat==1), 0,
        ifelse((!is.na(h_egoalt$mother_pidp)), -9, NA))))))),
  
  label = "Whether living with biological mother at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated_m at next wave" = 1,
             "possibly separated_m at next wave" = 2,
             "mother not alive at next wave" = 3))

#wave 8 leave: not live with fathers
h_egoalt$separated_f <- labelled(
  ifelse((!is.na(h_egoalt$father_pidp) & h_egoalt$h_enwstat==5), -8, 
  ifelse((!is.na(h_egoalt$father_pidp)  & h_egoalt$h_anwstat==5), 3,
  ifelse((!is.na(h_egoalt$father_pidp)  & 
          h_egoalt$h_anwstat==2 & h_egoalt$h_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(h_egoalt$father_pidp) & 
          h_egoalt$h_anwstat %in% c(3,4) & h_egoalt$h_enwstat==2), 2,
  ifelse((!is.na(h_egoalt$father_pidp) &
          h_egoalt$h_anwstat==2 & h_egoalt$h_enwstat==2), 1,
  ifelse((!is.na(h_egoalt$father_pidp) & h_egoalt$h_anwstat==1), 0,
        ifelse((!is.na(h_egoalt$father_pidp)), -9, NA))))))),
  
  label = "Whether living with biological father at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated at next wave" = 1,
             "possibly separated at next wave" = 2,
             "father not alive at next wave" = 3))


#fill the separate indicator
h_egoalt<-h_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_m,.direction="updown")
h_egoalt<-h_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_f,.direction="updown")

#find out who separate from both parents
#wave 9 leave: not live with both parents
h_egoalt$separated<- labelled(
  ifelse((h_egoalt$separated_m%in%c(1)&h_egoalt$separated_f%in%c(1)),1,
         
  ifelse((h_egoalt$separated_m%in%c(1)& is.na(h_egoalt$separated_f)),1,#one
  ifelse((h_egoalt$separated_f%in%c(1)& is.na(h_egoalt$separated_m)),1,
         
  ifelse((h_egoalt$separated_f%in%c(1)& h_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((h_egoalt$separated_m%in%c(1)& h_egoalt$separated_f%in%c(3)),1,      

         
  ifelse((h_egoalt$separated_m%in%c(2)&h_egoalt$separated_f%in%c(2)),2,
  ifelse((h_egoalt$separated_m%in%c(1)&h_egoalt$separated_f%in%c(2)),2,
  ifelse((h_egoalt$separated_m%in%c(2)&h_egoalt$separated_f%in%c(1)),2,
         
  ifelse((h_egoalt$separated_m%in%c(2)& is.na(h_egoalt$separated_f)),2,#one
  ifelse((h_egoalt$separated_f%in%c(2)& is.na(h_egoalt$separated_m)),2, 
         
  ifelse((h_egoalt$separated_f%in%c(2)& h_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((h_egoalt$separated_m%in%c(2)& h_egoalt$separated_f%in%c(3)),1,             
         
         
  ifelse((h_egoalt$separated_m%in%c(0)|h_egoalt$separated_f%in%c(0)),0,
  ifelse((h_egoalt$separated_m%in%c(-8)|h_egoalt$separated_f%in%c(-8)),-8,
         ifelse((!is.na(h_egoalt$separated_m)|!is.na(h_egoalt$separated_f)),-9,NA))))))))))))))),
         
  label = "Whether living with at least one biological parent at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together with at least one parent at next wave" = 0,
             "separated from both parents at next wave" = 1,
             "possibly separated from both parents at next wave" = 2))


#living with both parents indicator 
h_egoalt<-h_egoalt %>%group_by(pidp) %>%tidyr::fill(mother_pidp,.direction="updown")
h_egoalt<-h_egoalt %>%group_by(pidp) %>%tidyr::fill(father_pidp,.direction="updown")


#flags for living with mother or father this wave
h_egoalt<-h_egoalt%>%group_by(pidp)%>%mutate(livewithmother=ifelse(!is.na(h_egoalt$mother_pidp),1,0))
h_egoalt<-h_egoalt%>%group_by(pidp)%>%mutate(livewithfather=ifelse(!is.na(h_egoalt$father_pidp),1,0))

#living with both parents this wave
h_egoalt$liveboth<-c(0)
h_egoalt$liveboth[h_egoalt$livewithmother==1 & h_egoalt$livewithfather==1]<-1


#a column for living with mother, father, or both this wave
h_egoalt$lvpa<-c(0)
h_egoalt$lvpa[h_egoalt$livewithmother==1]<-1
h_egoalt$lvpa[h_egoalt$livewithfather==1]<-2
h_egoalt$lvpa[h_egoalt$liveboth==1]<-3

h_egoalt$lvpa<-factor(h_egoalt$lvpa, levels=c(0,1,2,3),labels=c("leaver","live with mother","live with father","live with both parents"))



```

```{r}
#merge datafiles
h_indresp<-h_indresp%>%dplyr::select(pidp, h_hidp, h_mastat_dv,
         h_age_dv, h_sex_dv, h_hiqual_dv,
         h_mastat_dv, h_racel_dv,h_fimnnet_dv,h_nchild_dv,h_urban_dv,h_gor_dv,h_jbstat,h_sf1,h_scsf1)


h_egoaltt<-h_egoalt%>%dplyr::select(pidp,h_hidp,mother_pidp,father_pidp,separated_m,separated_f,separated,lvpa)

#child levels
h_indrespc<-join_all(list(h_indresp,h_egoaltt), by='pidp', type='left')
h_indrespc<-h_indrespc %>% distinct_at(vars(pidp), .keep_all = TRUE)

h_indrespc<-h_indrespc[(which(!is.na(h_indrespc$mother_pidp)|!is.na(h_indrespc$father_pidp))),]



#add parent levels

h_indresppt <-h_indresp

#join parental data by pidp
father<-as.data.frame(h_indrespc$father_pidp)
mother<-as.data.frame(h_indrespc$mother_pidp)

colnames(father)[1]<-"pidp"
colnames(mother)[1]<-"pidp"

#attached parental values
h_indresppf<-join_all(list(father,h_indresppt), by='pidp', type='left')
h_indresppm<-join_all(list(mother,h_indresppt), by='pidp', type='left')

#for housing tenure
h_indresppf<-join_all(list(h_indresppf,h_hhresp), by='h_hidp', type='left')
h_indresppm<-join_all(list(h_indresppm,h_hhresp), by='h_hidp', type='left')


#rename var to differentiate parents
h_indresppf <-
        rename_at(h_indresppf, vars(starts_with("h_")), ~str_replace(., "h_", "ip_")) 
h_indresppm <-
        rename_at(h_indresppm, vars(starts_with("h_")), ~str_replace(., "h_", "im_")) 

#join to the child-level data by parent's pidp
colnames(h_indresppf)[1]<-"father_pidp"
colnames(h_indresppm)[1]<-"mother_pidp"

h_indresppf<-join_all(list(h_indrespc,h_indresppf), by="father_pidp", type='left')
h_indresppm<-join_all(list(h_indrespc,h_indresppm), by="mother_pidp", type='left')

h_indrespp<-join_all(list(h_indresppm,h_indresppf), by="pidp", type='left')

h_indrespp<-h_indrespp %>% distinct_at(vars(pidp), .keep_all = TRUE)

rm(h_egoaltt,h_indrespc,h_indresppf,h_indresppm,h_indresppt)

```

```{r}
## add cross wave identifier
h_indrespp$from_wave_h<-1
```

##### Wave 9

```{r}
mothers <- i_egoalt %>%
  filter(i_relationship_dv %in% c(4) & i_asex==2) %>%
  select(pidp, apidp)
mothers$mother_pidp=mothers$apidp

i_egoalt <- i_egoalt %>%
  full_join(mothers, by=c("pidp", "apidp")) 

rm(mothers)

fathers <- i_egoalt %>%
  filter(i_relationship_dv %in% c(4) & i_asex==1) %>%
  select(pidp, apidp)
fathers$father_pidp=fathers$apidp

i_egoalt <- i_egoalt %>%
  full_join(fathers, by=c("pidp", "apidp"))

rm(fathers)


#wave 9 leave: not live with mothers
i_egoalt$separated_m <- labelled(
  ifelse((!is.na(i_egoalt$mother_pidp) & i_egoalt$i_enwstat==5), -8, 
  ifelse((!is.na(i_egoalt$mother_pidp)  & i_egoalt$i_anwstat==5), 3,
  ifelse((!is.na(i_egoalt$mother_pidp)  & 
          i_egoalt$i_anwstat==2 & i_egoalt$i_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(i_egoalt$mother_pidp) & 
          i_egoalt$i_anwstat %in% c(3,4) & i_egoalt$i_enwstat==2), 2,
  ifelse((!is.na(i_egoalt$mother_pidp) &
          i_egoalt$i_anwstat==2 & i_egoalt$i_enwstat==2), 1,
  ifelse((!is.na(i_egoalt$mother_pidp)& i_egoalt$i_anwstat==1), 0,
        ifelse((!is.na(i_egoalt$mother_pidp)), -9, NA))))))),
  
  label = "Whether living with biological mother at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated_m at next wave" = 1,
             "possibly separated_m at next wave" = 2,
             "mother not alive at next wave" = 3))

#wave 9 leave: not live with fathers
i_egoalt$separated_f <- labelled(
  ifelse((!is.na(i_egoalt$father_pidp) & i_egoalt$i_enwstat==5), -8, 
  ifelse((!is.na(i_egoalt$father_pidp)  & i_egoalt$i_anwstat==5), 3,
  ifelse((!is.na(i_egoalt$father_pidp)  & 
          i_egoalt$i_anwstat==2 & i_egoalt$i_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(i_egoalt$father_pidp) & 
          i_egoalt$i_anwstat %in% c(3,4) & i_egoalt$i_enwstat==2), 2,
  ifelse((!is.na(i_egoalt$father_pidp) &
          i_egoalt$i_anwstat==2 & i_egoalt$i_enwstat==2), 1,
  ifelse((!is.na(i_egoalt$father_pidp) & i_egoalt$i_anwstat==1), 0,
        ifelse((!is.na(i_egoalt$father_pidp)), -9, NA))))))),
  
  label = "Whether living with biological father at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated at next wave" = 1,
             "possibly separated at next wave" = 2,
             "father not alive at next wave" = 3))


#fill the separate indicator
i_egoalt<-i_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_m,.direction="updown")
i_egoalt<-i_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_f,.direction="updown")

#find out who separate from both parents
#wave 9 leave: not live with both parents
i_egoalt$separated<- labelled(
  ifelse((i_egoalt$separated_m%in%c(1)&i_egoalt$separated_f%in%c(1)),1,
         
  ifelse((i_egoalt$separated_m%in%c(1)& is.na(i_egoalt$separated_f)),1,#one
  ifelse((i_egoalt$separated_f%in%c(1)& is.na(i_egoalt$separated_m)),1,
         
  ifelse((i_egoalt$separated_f%in%c(1)& i_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((i_egoalt$separated_m%in%c(1)& i_egoalt$separated_f%in%c(3)),1,      

         
  ifelse((i_egoalt$separated_m%in%c(2)&i_egoalt$separated_f%in%c(2)),2,
  ifelse((i_egoalt$separated_m%in%c(1)&i_egoalt$separated_f%in%c(2)),2,
  ifelse((i_egoalt$separated_m%in%c(2)&i_egoalt$separated_f%in%c(1)),2,
         
  ifelse((i_egoalt$separated_m%in%c(2)& is.na(i_egoalt$separated_f)),2,#one
  ifelse((i_egoalt$separated_f%in%c(2)& is.na(i_egoalt$separated_m)),2, 
         
  ifelse((i_egoalt$separated_f%in%c(2)& i_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((i_egoalt$separated_m%in%c(2)& i_egoalt$separated_f%in%c(3)),1,             
         
         
  ifelse((i_egoalt$separated_m%in%c(0)|i_egoalt$separated_f%in%c(0)),0,
  ifelse((i_egoalt$separated_m%in%c(-8)|i_egoalt$separated_f%in%c(-8)),-8,
         ifelse((!is.na(i_egoalt$separated_m)|!is.na(i_egoalt$separated_f)),-9,NA))))))))))))))),
         
  label = "Whether living with at least one biological parent at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together with at least one parent at next wave" = 0,
             "separated from both parents at next wave" = 1,
             "possibly separated from both parents at next wave" = 2))


#living with both parents indicator 
i_egoalt<-i_egoalt %>%group_by(pidp) %>%tidyr::fill(mother_pidp,.direction="updown")
i_egoalt<-i_egoalt %>%group_by(pidp) %>%tidyr::fill(father_pidp,.direction="updown")


#flags for living with mother or father this wave
i_egoalt<-i_egoalt%>%group_by(pidp)%>%mutate(livewithmother=ifelse(!is.na(i_egoalt$mother_pidp),1,0))
i_egoalt<-i_egoalt%>%group_by(pidp)%>%mutate(livewithfather=ifelse(!is.na(i_egoalt$father_pidp),1,0))

#living with both parents this wave
i_egoalt$liveboth<-c(0)
i_egoalt$liveboth[i_egoalt$livewithmother==1 & i_egoalt$livewithfather==1]<-1


#a column for living with mother, father, or both this wave
i_egoalt$lvpa<-c(0)
i_egoalt$lvpa[i_egoalt$livewithmother==1]<-1
i_egoalt$lvpa[i_egoalt$livewithfather==1]<-2
i_egoalt$lvpa[i_egoalt$liveboth==1]<-3

i_egoalt$lvpa<-factor(i_egoalt$lvpa, levels=c(0,1,2,3),labels=c("leaver","live with mother","live with father","live with both parents"))


```

```{r}
#merge datafiles
i_indresp<-i_indresp%>%dplyr::select(pidp, i_hidp, i_mastat_dv,
         i_age_dv, i_sex_dv, i_hiqual_dv,
         i_mastat_dv, i_racel_dv,i_fimnnet_dv,i_nchild_dv,i_urban_dv,i_gor_dv,i_jbstat,i_sf1,i_scsf1)


i_egoaltt<-i_egoalt%>%dplyr::select(pidp,i_hidp,mother_pidp,father_pidp,separated_m,separated_f,separated,lvpa)

#child levels
i_indrespc<-join_all(list(i_indresp,i_egoaltt), by='pidp', type='left')
i_indrespc<-i_indrespc %>% distinct_at(vars(pidp), .keep_all = TRUE)

i_indrespc<-i_indrespc[(which(!is.na(i_indrespc$mother_pidp)|!is.na(i_indrespc$father_pidp))),]



#add parent levels

i_indresppt <-i_indresp

#join parental data by pidp
father<-as.data.frame(i_indrespc$father_pidp)
mother<-as.data.frame(i_indrespc$mother_pidp)

colnames(father)[1]<-"pidp"
colnames(mother)[1]<-"pidp"

#attached parental values
i_indresppf<-join_all(list(father,i_indresppt), by='pidp', type='left')
i_indresppm<-join_all(list(mother,i_indresppt), by='pidp', type='left')

#for housing tenure
i_indresppf<-join_all(list(i_indresppf,i_hhresp), by='i_hidp', type='left')
i_indresppm<-join_all(list(i_indresppm,i_hhresp), by='i_hidp', type='left')


#rename var to differentiate parents
i_indresppf <-
        rename_at(i_indresppf, vars(starts_with("i_")), ~str_replace(., "i_", "ip_")) 
i_indresppm <-
        rename_at(i_indresppm, vars(starts_with("i_")), ~str_replace(., "i_", "im_")) 

#join to the child-level data by parent's pidp
colnames(i_indresppf)[1]<-"father_pidp"
colnames(i_indresppm)[1]<-"mother_pidp"

i_indresppf<-join_all(list(i_indrespc,i_indresppf), by="father_pidp", type='left')
i_indresppm<-join_all(list(i_indrespc,i_indresppm), by="mother_pidp", type='left')

i_indrespp<-join_all(list(i_indresppm,i_indresppf), by="pidp", type='left')

i_indrespp<-i_indrespp %>% distinct_at(vars(pidp), .keep_all = TRUE)

rm(i_egoaltt,i_indrespc,i_indresppf,i_indresppm,i_indresppt)
```

```{r}
## add cross wave identifier
i_indrespp$from_wave_i<-1
```

##### Wave 10

```{r}
mothers <- j_egoalt %>%
  filter(j_relationship_dv %in% c(4) & j_asex==2) %>%
  select(pidp, apidp)
mothers$mother_pidp=mothers$apidp

j_egoalt <- j_egoalt %>%
  full_join(mothers, by=c("pidp", "apidp")) 

rm(mothers)

fathers <- j_egoalt %>%
  filter(j_relationship_dv %in% c(4) & j_asex==1) %>%
  select(pidp, apidp)
fathers$father_pidp=fathers$apidp

j_egoalt <- j_egoalt %>%
  full_join(fathers, by=c("pidp", "apidp"))

rm(fathers)


#wave 9 leave: not live with mothers
j_egoalt$separated_m <- labelled(
  ifelse((!is.na(j_egoalt$mother_pidp) & j_egoalt$j_enwstat==5), -8, 
  ifelse((!is.na(j_egoalt$mother_pidp)  & j_egoalt$j_anwstat==5), 3,
  ifelse((!is.na(j_egoalt$mother_pidp)  & 
          j_egoalt$j_anwstat==2 & j_egoalt$j_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(j_egoalt$mother_pidp) & 
          j_egoalt$j_anwstat %in% c(3,4) & j_egoalt$j_enwstat==2), 2,
  ifelse((!is.na(j_egoalt$mother_pidp) &
          j_egoalt$j_anwstat==2 & j_egoalt$j_enwstat==2), 1,
  ifelse((!is.na(j_egoalt$mother_pidp)& j_egoalt$j_anwstat==1), 0,
        ifelse((!is.na(j_egoalt$mother_pidp)), -9, NA))))))),
  
  label = "Whether living with biological mother at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated_m at next wave" = 1,
             "possibly separated_m at next wave" = 2,
             "mother not alive at next wave" = 3))

#wave 9 leave: not live with fathers
j_egoalt$separated_f <- labelled(
  ifelse((!is.na(j_egoalt$father_pidp) & j_egoalt$j_enwstat==5), -8, 
  ifelse((!is.na(j_egoalt$father_pidp)  & j_egoalt$j_anwstat==5), 3,
  ifelse((!is.na(j_egoalt$father_pidp)  & 
          j_egoalt$j_anwstat==2 & j_egoalt$j_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(j_egoalt$father_pidp) & 
          j_egoalt$j_anwstat %in% c(3,4) & j_egoalt$j_enwstat==2), 2,
  ifelse((!is.na(j_egoalt$father_pidp) &
          j_egoalt$j_anwstat==2 & j_egoalt$j_enwstat==2), 1,
  ifelse((!is.na(j_egoalt$father_pidp) & j_egoalt$j_anwstat==1), 0,
        ifelse((!is.na(j_egoalt$father_pidp)), -9, NA))))))),
  
  label = "Whether living with biological father at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated at next wave" = 1,
             "possibly separated at next wave" = 2,
             "father not alive at next wave" = 3))


#fill the separate indicator
j_egoalt<-j_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_m,.direction="updown")
j_egoalt<-j_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_f,.direction="updown")

#find out who separate from both parents
#wave 9 leave: not live with both parents
j_egoalt$separated<- labelled(
  ifelse((j_egoalt$separated_m%in%c(1)&j_egoalt$separated_f%in%c(1)),1,
         
  ifelse((j_egoalt$separated_m%in%c(1)& is.na(j_egoalt$separated_f)),1,#one
  ifelse((j_egoalt$separated_f%in%c(1)& is.na(j_egoalt$separated_m)),1,
         
  ifelse((j_egoalt$separated_f%in%c(1)& j_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((j_egoalt$separated_m%in%c(1)& j_egoalt$separated_f%in%c(3)),1,      

         
  ifelse((j_egoalt$separated_m%in%c(2)&j_egoalt$separated_f%in%c(2)),2,
  ifelse((j_egoalt$separated_m%in%c(1)&j_egoalt$separated_f%in%c(2)),2,
  ifelse((j_egoalt$separated_m%in%c(2)&j_egoalt$separated_f%in%c(1)),2,
         
  ifelse((j_egoalt$separated_m%in%c(2)& is.na(j_egoalt$separated_f)),2,#one
  ifelse((j_egoalt$separated_f%in%c(2)& is.na(j_egoalt$separated_m)),2, 
         
  ifelse((j_egoalt$separated_f%in%c(2)& j_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((j_egoalt$separated_m%in%c(2)& j_egoalt$separated_f%in%c(3)),1,             
         
         
  ifelse((j_egoalt$separated_m%in%c(0)|j_egoalt$separated_f%in%c(0)),0,
  ifelse((j_egoalt$separated_m%in%c(-8)|j_egoalt$separated_f%in%c(-8)),-8,
         ifelse((!is.na(j_egoalt$separated_m)|!is.na(j_egoalt$separated_f)),-9,NA))))))))))))))),
         
  label = "Whether living with at least one biological parent at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together with at least one parent at next wave" = 0,
             "separated from both parents at next wave" = 1,
             "possibly separated from both parents at next wave" = 2))


#living with both parents indicator 
j_egoalt<-j_egoalt %>%group_by(pidp) %>%tidyr::fill(mother_pidp,.direction="updown")
j_egoalt<-j_egoalt %>%group_by(pidp) %>%tidyr::fill(father_pidp,.direction="updown")


#flags for living with mother or father this wave
j_egoalt<-j_egoalt%>%group_by(pidp)%>%mutate(livewithmother=ifelse(!is.na(j_egoalt$mother_pidp),1,0))
j_egoalt<-j_egoalt%>%group_by(pidp)%>%mutate(livewithfather=ifelse(!is.na(j_egoalt$father_pidp),1,0))

#living with both parents this wave
j_egoalt$liveboth<-c(0)
j_egoalt$liveboth[j_egoalt$livewithmother==1 & j_egoalt$livewithfather==1]<-1


#a column for living with mother, father, or both this wave
j_egoalt$lvpa<-c(0)
j_egoalt$lvpa[j_egoalt$livewithmother==1]<-1
j_egoalt$lvpa[j_egoalt$livewithfather==1]<-2
j_egoalt$lvpa[j_egoalt$liveboth==1]<-3

j_egoalt$lvpa<-factor(j_egoalt$lvpa, levels=c(0,1,2,3),labels=c("leaver","live with mother","live with father","live with both parents"))

```

```{r}
#merge datafiles
j_indresp<-j_indresp%>%dplyr::select(pidp, j_hidp, j_mastat_dv,
         j_age_dv, j_sex_dv, j_hiqual_dv,
         j_mastat_dv, j_racel_dv,j_fimnnet_dv,j_nchild_dv,j_urban_dv,j_gor_dv,j_jbstat,j_sf1,j_scsf1)


j_egoaltt<-j_egoalt%>%dplyr::select(pidp,j_hidp,mother_pidp,father_pidp,separated_m,separated_f,separated,lvpa)

#child levels
j_indrespc<-join_all(list(j_indresp,j_egoaltt), by='pidp', type='left')
j_indrespc<-j_indrespc %>% distinct_at(vars(pidp), .keep_all = TRUE)

j_indrespc<-j_indrespc[(which(!is.na(j_indrespc$mother_pidp)|!is.na(j_indrespc$father_pidp))),]



#add parent levels

j_indresppt <-j_indresp

#join parental data by pidp
father<-as.data.frame(j_indrespc$father_pidp)
mother<-as.data.frame(j_indrespc$mother_pidp)

colnames(father)[1]<-"pidp"
colnames(mother)[1]<-"pidp"

#attached parental values
j_indresppf<-join_all(list(father,j_indresppt), by='pidp', type='left')
j_indresppm<-join_all(list(mother,j_indresppt), by='pidp', type='left')

#for housing tenure
j_indresppf<-join_all(list(j_indresppf,j_hhresp), by='j_hidp', type='left')
j_indresppm<-join_all(list(j_indresppm,j_hhresp), by='j_hidp', type='left')


#rename var to differentiate parents
j_indresppf <-
        rename_at(j_indresppf, vars(starts_with("j_")), ~str_replace(., "j_", "ip_")) 
j_indresppm <-
        rename_at(j_indresppm, vars(starts_with("j_")), ~str_replace(., "j_", "im_")) 

#join to the child-level data by parent's pidp
colnames(j_indresppf)[1]<-"father_pidp"
colnames(j_indresppm)[1]<-"mother_pidp"

j_indresppf<-join_all(list(j_indrespc,j_indresppf), by="father_pidp", type='left')
j_indresppm<-join_all(list(j_indrespc,j_indresppm), by="mother_pidp", type='left')

j_indrespp<-join_all(list(j_indresppm,j_indresppf), by="pidp", type='left')

j_indrespp<-j_indrespp %>% distinct_at(vars(pidp), .keep_all = TRUE)

rm(j_egoaltt,j_indrespc,j_indresppf,j_indresppm,j_indresppt)

```

```{r}
j_indrespp$from_wave_j<-1
```

##### Wave 11

```{r}
mothers <- k_egoalt %>%
  filter(k_relationship_dv %in% c(4) & k_asex==2) %>%
  select(pidp, apidp)
mothers$mother_pidp=mothers$apidp

k_egoalt <- k_egoalt %>%
  full_join(mothers, by=c("pidp", "apidp")) 

rm(mothers)

fathers <- k_egoalt %>%
  filter(k_relationship_dv %in% c(4) & k_asex==1) %>%
  select(pidp, apidp)
fathers$father_pidp=fathers$apidp

k_egoalt <- k_egoalt %>%
  full_join(fathers, by=c("pidp", "apidp"))

rm(fathers)


#wave 8 leave: not live with mothers
k_egoalt$separated_m <- labelled(
  ifelse((!is.na(k_egoalt$mother_pidp) & k_egoalt$k_enwstat==5), -8, 
  ifelse((!is.na(k_egoalt$mother_pidp)  & k_egoalt$k_anwstat==5), 3,
  ifelse((!is.na(k_egoalt$mother_pidp)  & 
          k_egoalt$k_anwstat==2 & k_egoalt$k_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(k_egoalt$mother_pidp) & 
          k_egoalt$k_anwstat %in% c(3,4) & k_egoalt$k_enwstat==2), 2,
  ifelse((!is.na(k_egoalt$mother_pidp) &
          k_egoalt$k_anwstat==2 & k_egoalt$k_enwstat==2), 1,
  ifelse((!is.na(k_egoalt$mother_pidp)& k_egoalt$k_anwstat==1), 0,
        ifelse((!is.na(k_egoalt$mother_pidp)), -9, NA))))))),
  
  label = "Whether living with biological mother at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated_m at next wave" = 1,
             "possibly separated_m at next wave" = 2,
             "mother not alive at next wave" = 3))

#wave 8 leave: not live with fathers
k_egoalt$separated_f <- labelled(
  ifelse((!is.na(k_egoalt$father_pidp) & k_egoalt$k_enwstat==5), -8, 
  ifelse((!is.na(k_egoalt$father_pidp)  & k_egoalt$k_anwstat==5), 3,
  ifelse((!is.na(k_egoalt$father_pidp)  & 
          k_egoalt$k_anwstat==2 & k_egoalt$k_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(k_egoalt$father_pidp) & 
          k_egoalt$k_anwstat %in% c(3,4) & k_egoalt$k_enwstat==2), 2,
  ifelse((!is.na(k_egoalt$father_pidp) &
          k_egoalt$k_anwstat==2 & k_egoalt$k_enwstat==2), 1,
  ifelse((!is.na(k_egoalt$father_pidp) & k_egoalt$k_anwstat==1), 0,
        ifelse((!is.na(k_egoalt$father_pidp)), -9, NA))))))),
  
  label = "Whether living with biological father at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated at next wave" = 1,
             "possibly separated at next wave" = 2,
             "father not alive at next wave" = 3))


#fill the separate indicator
k_egoalt<-k_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_m,.direction="updown")
k_egoalt<-k_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_f,.direction="updown")

#find out who separate from both parents
#wave 9 leave: not live with both parents
k_egoalt$separated<- labelled(
  ifelse((k_egoalt$separated_m%in%c(1)&k_egoalt$separated_f%in%c(1)),1,
         
  ifelse((k_egoalt$separated_m%in%c(1)& is.na(k_egoalt$separated_f)),1,#one
  ifelse((k_egoalt$separated_f%in%c(1)& is.na(k_egoalt$separated_m)),1,
         
  ifelse((k_egoalt$separated_f%in%c(1)& k_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((k_egoalt$separated_m%in%c(1)& k_egoalt$separated_f%in%c(3)),1,      

         
  ifelse((k_egoalt$separated_m%in%c(2)&k_egoalt$separated_f%in%c(2)),2,
  ifelse((k_egoalt$separated_m%in%c(1)&k_egoalt$separated_f%in%c(2)),2,
  ifelse((k_egoalt$separated_m%in%c(2)&k_egoalt$separated_f%in%c(1)),2,
         
  ifelse((k_egoalt$separated_m%in%c(2)& is.na(k_egoalt$separated_f)),2,#one
  ifelse((k_egoalt$separated_f%in%c(2)& is.na(k_egoalt$separated_m)),2, 
         
  ifelse((k_egoalt$separated_f%in%c(2)& k_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((k_egoalt$separated_m%in%c(2)& k_egoalt$separated_f%in%c(3)),1,             
         
         
  ifelse((k_egoalt$separated_m%in%c(0)|k_egoalt$separated_f%in%c(0)),0,
  ifelse((k_egoalt$separated_m%in%c(-8)|k_egoalt$separated_f%in%c(-8)),-8,
         ifelse((!is.na(k_egoalt$separated_m)|!is.na(k_egoalt$separated_f)),-9,NA))))))))))))))),
         
  label = "Whether living with at least one biological parent at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together with at least one parent at next wave" = 0,
             "separated from both parents at next wave" = 1,
             "possibly separated from both parents at next wave" = 2))


#living with both parents indicator 
k_egoalt<-k_egoalt %>%group_by(pidp) %>%tidyr::fill(mother_pidp,.direction="updown")
k_egoalt<-k_egoalt %>%group_by(pidp) %>%tidyr::fill(father_pidp,.direction="updown")


#flags for living with mother or father this wave
k_egoalt<-k_egoalt%>%group_by(pidp)%>%mutate(livewithmother=ifelse(!is.na(k_egoalt$mother_pidp),1,0))
k_egoalt<-k_egoalt%>%group_by(pidp)%>%mutate(livewithfather=ifelse(!is.na(k_egoalt$father_pidp),1,0))

#living with both parents this wave
k_egoalt$liveboth<-c(0)
k_egoalt$liveboth[k_egoalt$livewithmother==1 & k_egoalt$livewithfather==1]<-1


#a column for living with mother, father, or both this wave
k_egoalt$lvpa<-c(0)
k_egoalt$lvpa[k_egoalt$livewithmother==1]<-1
k_egoalt$lvpa[k_egoalt$livewithfather==1]<-2
k_egoalt$lvpa[k_egoalt$liveboth==1]<-3

k_egoalt$lvpa<-factor(k_egoalt$lvpa, levels=c(0,1,2,3),labels=c("leaver","live with mother","live with father","live with both parents"))

```

```{r}
#merge datafiles
k_indresp<-k_indresp%>%dplyr::select(pidp, k_hidp, k_mastat_dv,
         k_age_dv, k_sex_dv, k_hiqual_dv,
         k_mastat_dv, k_racel_dv,k_fimnnet_dv,k_nchild_dv,k_urban_dv,k_gor_dv,k_jbstat,k_sf1,k_scsf1)


k_egoaltt<-k_egoalt%>%dplyr::select(pidp,k_hidp,mother_pidp,father_pidp,separated_m,separated_f,separated,lvpa)

#child levels
k_indrespc<-join_all(list(k_indresp,k_egoaltt), by='pidp', type='left')
k_indrespc<-k_indrespc %>% distinct_at(vars(pidp), .keep_all = TRUE)

k_indrespc<-k_indrespc[(which(!is.na(k_indrespc$mother_pidp)|!is.na(k_indrespc$father_pidp))),]



#add parent levels

k_indresppt <-k_indresp

#join parental data by pidp
father<-as.data.frame(k_indrespc$father_pidp)
mother<-as.data.frame(k_indrespc$mother_pidp)

colnames(father)[1]<-"pidp"
colnames(mother)[1]<-"pidp"

#attached parental values
k_indresppf<-join_all(list(father,k_indresppt), by='pidp', type='left')
k_indresppm<-join_all(list(mother,k_indresppt), by='pidp', type='left')

#for housing tenure
k_indresppf<-join_all(list(k_indresppf,k_hhresp), by='k_hidp', type='left')
k_indresppm<-join_all(list(k_indresppm,k_hhresp), by='k_hidp', type='left')


#rename var to differentiate parents
k_indresppf <-
        rename_at(k_indresppf, vars(starts_with("k_")), ~str_replace(., "k_", "ip_")) 
k_indresppm <-
        rename_at(k_indresppm, vars(starts_with("k_")), ~str_replace(., "k_", "im_")) 

#join to the child-level data by parent's pidp
colnames(k_indresppf)[1]<-"father_pidp"
colnames(k_indresppm)[1]<-"mother_pidp"

k_indresppf<-join_all(list(k_indrespc,k_indresppf), by="father_pidp", type='left')
k_indresppm<-join_all(list(k_indrespc,k_indresppm), by="mother_pidp", type='left')

k_indrespp<-join_all(list(k_indresppm,k_indresppf), by="pidp", type='left')

k_indrespp<-k_indrespp %>% distinct_at(vars(pidp), .keep_all = TRUE)

rm(k_egoaltt,k_indrespc,k_indresppf,k_indresppm,k_indresppt)

```

```{r}
k_indrespp$from_wave_k<-1
```

##### Wave 12

```{r}
mothers <- l_egoalt %>%
  filter(l_relationship_dv %in% c(4) & l_asex==2) %>%
  select(pidp, apidp)
mothers$mother_pidp=mothers$apidp

l_egoalt <- l_egoalt %>%
  full_join(mothers, by=c("pidp", "apidp")) 

rm(mothers)

fathers <- l_egoalt %>%
  filter(l_relationship_dv %in% c(4) & l_asex==1) %>%
  select(pidp, apidp)
fathers$father_pidp=fathers$apidp

l_egoalt <- l_egoalt %>%
  full_join(fathers, by=c("pidp", "apidp"))

rm(fathers)


#wave 8 leave: not live with mothers
l_egoalt$separated_m <- labelled(
  ifelse((!is.na(l_egoalt$mother_pidp) & l_egoalt$l_enwstat==5), -8, 
  ifelse((!is.na(l_egoalt$mother_pidp)  & l_egoalt$l_anwstat==5), 3,
  ifelse((!is.na(l_egoalt$mother_pidp)  & 
          l_egoalt$l_anwstat==2 & l_egoalt$l_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(l_egoalt$mother_pidp) & 
          l_egoalt$l_anwstat %in% c(3,4) & l_egoalt$l_enwstat==2), 2,
  ifelse((!is.na(l_egoalt$mother_pidp) &
          l_egoalt$l_anwstat==2 & l_egoalt$l_enwstat==2), 1,
  ifelse((!is.na(l_egoalt$mother_pidp)& l_egoalt$l_anwstat==1), 0,
        ifelse((!is.na(l_egoalt$mother_pidp)), -9, NA))))))),
  
  label = "Whether living with biological mother at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated_m at next wave" = 1,
             "possibly separated_m at next wave" = 2,
             "mother not alive at next wave" = 3))

#wave 8 leave: not live with fathers
l_egoalt$separated_f <- labelled(
  ifelse((!is.na(l_egoalt$father_pidp) & l_egoalt$l_enwstat==5), -8, 
  ifelse((!is.na(l_egoalt$father_pidp)  & l_egoalt$l_anwstat==5), 3,
  ifelse((!is.na(l_egoalt$father_pidp)  & 
          l_egoalt$l_anwstat==2 & l_egoalt$l_enwstat %in% c(3,4)), 2,
  ifelse((!is.na(l_egoalt$father_pidp) & 
          l_egoalt$l_anwstat %in% c(3,4) & l_egoalt$l_enwstat==2), 2,
  ifelse((!is.na(l_egoalt$father_pidp) &
          l_egoalt$l_anwstat==2 & l_egoalt$l_enwstat==2), 1,
  ifelse((!is.na(l_egoalt$father_pidp) & l_egoalt$l_anwstat==1), 0,
        ifelse((!is.na(l_egoalt$father_pidp)), -9, NA))))))),
  
  label = "Whether living with biological father at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together at next wave" = 0, "separated at next wave" = 1,
             "possibly separated at next wave" = 2,
             "father not alive at next wave" = 3))


#fill the separate indicator
l_egoalt<-l_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_m,.direction="updown")
l_egoalt<-l_egoalt %>%group_by(pidp) %>%tidyr::fill(separated_f,.direction="updown")

#find out who separate from both parents
#wave 9 leave: not live with both parents
l_egoalt$separated<- labelled(
  ifelse((l_egoalt$separated_m%in%c(1)&l_egoalt$separated_f%in%c(1)),1,
         
  ifelse((l_egoalt$separated_m%in%c(1)& is.na(l_egoalt$separated_f)),1,#one
  ifelse((l_egoalt$separated_f%in%c(1)& is.na(l_egoalt$separated_m)),1,
         
  ifelse((l_egoalt$separated_f%in%c(1)& l_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((l_egoalt$separated_m%in%c(1)& l_egoalt$separated_f%in%c(3)),1,      

         
  ifelse((l_egoalt$separated_m%in%c(2)&l_egoalt$separated_f%in%c(2)),2,
  ifelse((l_egoalt$separated_m%in%c(1)&l_egoalt$separated_f%in%c(2)),2,
  ifelse((l_egoalt$separated_m%in%c(2)&l_egoalt$separated_f%in%c(1)),2,
         
  ifelse((l_egoalt$separated_m%in%c(2)& is.na(l_egoalt$separated_f)),2,#one
  ifelse((l_egoalt$separated_f%in%c(2)& is.na(l_egoalt$separated_m)),2, 
         
  ifelse((l_egoalt$separated_f%in%c(2)& l_egoalt$separated_m%in%c(3)),1,#deadth
  ifelse((l_egoalt$separated_m%in%c(2)& l_egoalt$separated_f%in%c(3)),1,             
         
         
  ifelse((l_egoalt$separated_m%in%c(0)|l_egoalt$separated_f%in%c(0)),0,
  ifelse((l_egoalt$separated_m%in%c(-8)|l_egoalt$separated_f%in%c(-8)),-8,
         ifelse((!is.na(l_egoalt$separated_m)|!is.na(l_egoalt$separated_f)),-9,NA))))))))))))))),
         
  label = "Whether living with at least one biological parent at the next wave",
  labels = c("missing" = -9, 
             "ego not alive at next wave" = -8,
             "living together with at least one parent at next wave" = 0,
             "separated from both parents at next wave" = 1,
             "possibly separated from both parents at next wave" = 2))


#living with both parents indicator 
l_egoalt<-l_egoalt %>%group_by(pidp) %>%tidyr::fill(mother_pidp,.direction="updown")
l_egoalt<-l_egoalt %>%group_by(pidp) %>%tidyr::fill(father_pidp,.direction="updown")


#flags for living with mother or father this wave
l_egoalt<-l_egoalt%>%group_by(pidp)%>%mutate(livewithmother=ifelse(!is.na(l_egoalt$mother_pidp),1,0))
l_egoalt<-l_egoalt%>%group_by(pidp)%>%mutate(livewithfather=ifelse(!is.na(l_egoalt$father_pidp),1,0))

#living with both parents this wave
l_egoalt$liveboth<-c(0)
l_egoalt$liveboth[l_egoalt$livewithmother==1 & l_egoalt$livewithfather==1]<-1


#a column for living with mother, father, or both this wave
l_egoalt$lvpa<-c(0)
l_egoalt$lvpa[l_egoalt$livewithmother==1]<-1
l_egoalt$lvpa[l_egoalt$livewithfather==1]<-2
l_egoalt$lvpa[l_egoalt$liveboth==1]<-3

l_egoalt$lvpa<-factor(l_egoalt$lvpa, levels=c(0,1,2,3),labels=c("leaver","live with mother","live with father","live with both parents"))

```

```{r}
#merge datafiles
l_indresp<-l_indresp%>%dplyr::select(pidp, l_hidp, l_mastat_dv,
         l_age_dv, l_sex_dv, l_hiqual_dv,
         l_mastat_dv, l_racel_dv,l_fimnnet_dv,l_nchild_dv,l_urban_dv,l_gor_dv,l_jbstat,l_sf1,l_scsf1)


l_egoaltt<-l_egoalt%>%dplyr::select(pidp,l_hidp,mother_pidp,father_pidp,separated_m,separated_f,separated,lvpa)

#child levels
l_indrespc<-join_all(list(l_indresp,l_egoaltt), by='pidp', type='left')
l_indrespc<-l_indrespc %>% distinct_at(vars(pidp), .keep_all = TRUE)

l_indrespc<-l_indrespc[(which(!is.na(l_indrespc$mother_pidp)|!is.na(l_indrespc$father_pidp))),]



#add parent levels

l_indresppt <-l_indresp

#join parental data by pidp
father<-as.data.frame(l_indrespc$father_pidp)
mother<-as.data.frame(l_indrespc$mother_pidp)

colnames(father)[1]<-"pidp"
colnames(mother)[1]<-"pidp"

#attached parental values
l_indresppf<-join_all(list(father,l_indresppt), by='pidp', type='left')
l_indresppm<-join_all(list(mother,l_indresppt), by='pidp', type='left')

#for housing tenure
l_indresppf<-join_all(list(l_indresppf,l_hhresp), by='l_hidp', type='left')
l_indresppm<-join_all(list(l_indresppm,l_hhresp), by='l_hidp', type='left')


#rename var to differentiate parents
l_indresppf <-
        rename_at(l_indresppf, vars(starts_with("l_")), ~str_replace(., "l_", "ip_")) 
l_indresppm <-
        rename_at(l_indresppm, vars(starts_with("l_")), ~str_replace(., "l_", "im_")) 

#join to the child-level data by parent's pidp
colnames(l_indresppf)[1]<-"father_pidp"
colnames(l_indresppm)[1]<-"mother_pidp"

l_indresppf<-join_all(list(l_indrespc,l_indresppf), by="father_pidp", type='left')
l_indresppm<-join_all(list(l_indrespc,l_indresppm), by="mother_pidp", type='left')

l_indrespp<-join_all(list(l_indresppm,l_indresppf), by="pidp", type='left')

l_indrespp<-l_indrespp %>% distinct_at(vars(pidp), .keep_all = TRUE)

rm(l_egoaltt,l_indrespc,l_indresppf,l_indresppm,l_indresppt)

```

```{r}
l_indrespp$from_wave_l<-1
```

#### Merging all

```{r}
#remove all the prefixes
h_indrespp <-
        rename_at(h_indrespp, vars(starts_with("h_")), ~str_replace(., "h_", ""))
i_indrespp <-
        rename_at(i_indrespp, vars(starts_with("i_")), ~str_replace(., "i_", ""))
j_indrespp <-
        rename_at(j_indrespp, vars(starts_with("j_")), ~str_replace(., "j_", ""))
k_indrespp <-
        rename_at(k_indrespp, vars(starts_with("k_")), ~str_replace(., "k_", ""))
l_indrespp <-
        rename_at(l_indrespp, vars(starts_with("l_")), ~str_replace(., "l_", ""))


indrespb<-rbind.fill(h_indrespp,i_indrespp,j_indrespp,k_indrespp,l_indrespp)

```

#### Coding variables

child level

```{r}
#child level

#level of education completed
indrespb$edu<-c(0)
indrespb$edu[indrespb$hiqual_dv%in%c(9)]<-1 #low (everything below)
indrespb$edu[indrespb$hiqual_dv%in%c(3,4,5)]<-2 #middle (completed A-level or secondary high school)
indrespb$edu[indrespb$hiqual_dv%in%c(1,2)]<-3 #high (have a degree)
indrespb$edu[indrespb$edu==0]<-NA

print(attr(indrespb$hiqual_dv,"labels"))
indrespb$edu<-factor(indrespb$edu,levels=c(1,2,3),labels=c('low',"middle","high"))

#two-category education
indrespb$edu2<-c(0)
indrespb$edu2[indrespb$hiqual_dv%in%c(1,2)]<-1 #high (have a degree)
indrespb$edu2<-factor(indrespb$edu2,levels=c(0,1),labels=c("have a degree","do not have a degree"))

#family structure
indrespb$family_status<-c(0)
#coresident partner and child
indrespb$family_status[indrespb$nchild_dv>=1 & indrespb$mastat_dv%in%c(2,3,10)]<-1
#coresident partner
indrespb$family_status[indrespb$mastat_dv%in%c(2,3,10)&indrespb$nchild_dv==0]<-2
#coresident child
indrespb$family_status[indrespb$nchild_dv>=1&indrespb$mastat_dv%in%c(0,1,4,5,6,7,8,9)]<-3
#no coresident partner or child=0

indrespb$family_status<-factor(indrespb$family_status,levels=c(0,1,2,3),labels=c("No coresident partner or child","coresident partner and child","coresident partner","coresident child"))

#age centered
indrespb$age_centered<-indrespb$age_dv-16
#age squared
indrespb$age_sq<-indrespb$age_centered**2


#logarithm of income
indrespb$fimnnet_dv[indrespb$fimnnet_dv<0]<-0
#add 1 to all income values to avoid -inf in log transformation
indrespb$incomeadd<-indrespb$fimnnet_dv+1
indrespb$lnincome<-log(indrespb$incomeadd)
indrespb<-indrespb%>%dplyr::select(-incomeadd)


#racel groups 
indrespb$racel<-c(0)
indrespb$racel[indrespb$racel_dv%in%c(1,2,3,4)]<-1 #Whites
indrespb$racel[indrespb$racel_dv%in%c(14,15)]<-2 #African Caribbean
indrespb$racel[indrespb$racel_dv%in%c(9,10,11)]<-3 #Indiani,Pakistani, Bangladeshi
indrespb$racel[indrespb$racel_dv%in%c(12,13)]<-4 #Other Asian
indrespb$racel[indrespb$racel==0]<-5 #Other

indrespb$racel<-factor(indrespb$racel,levels=c(1,2,3,4,5),labels=c("Whites","African Caribbean","Indiani,Pakistani, Bangladeshi","Other Asian","Other"))

indrespb$racel[is.na(indrespb$racel_dv)]<-NA

##two-categories
indrespb$racel2<-c(0)
indrespb$racel2[indrespb$racel_dv%in%c(1,2,3,4)]<-1 #Whites
indrespb$racel2[!(indrespb$racel_dv%in%c(1,2,3,4))]<-2 #non-Whites
indrespb$racel2[is.na(indrespb$racel_dv)]<-NA

indrespb$racel2<-factor(indrespb$racel2,levels=c(1,2),labels=c("Whites","non-Whites"))

#education status
indrespb$student<-c(0)
indrespb$student[indrespb$jbstat%in%c(7)]<-1 #full-time student
indrespb$student[is.na(indrespb$jbstat)]<-NA

#employment status
indrespb$unemployed<-c(0)
indrespb$unemployed[indrespb$jbstat%in%c(3)]<-1 
indrespb$unemployed[is.na(indrespb$jbstat)]<-NA

#housing prices
indrespb$houseprices<-c(0)
indrespb$houseprices[indrespb$gor_dv%in%c(1,10,11,12)]<-1 #low
indrespb$houseprices[indrespb$gor_dv%in%c(2,3,4,5)]<-2 #middle
indrespb$houseprices[indrespb$gor_dv%in%c(6,7,8,9)]<-3 #high
indrespb$houseprices[is.na(indrespb$gor_dv)]<-NA

indrespb$houseprices<-factor(indrespb$houseprices,levels=c(1,2,3),labels=c("low","middle","high"))

#urban,sex binary
indrespb$sex<-c(0)
indrespb$sex[indrespb$sex_dv==2]<-1 #females
indrespb$sex[is.na(indrespb$sex_dv)]<-NA

indrespb$urban<-c(0)
indrespb$urban[indrespb$urban_dv==1]<-1 #urban areas
indrespb$urban[is.na(indrespb$urban_dv)]<-NA

indrespb<-indrespb%>%mutate(health=coalesce(sf1,scsf1))

#living with both biological parents
indrespb$liveboth<-c(0)
indrespb$liveboth[indrespb$lvpa=="live with both parents"]<-1
indrespb$liveboth<-factor(indrespb$liveboth,levels=c(0,1),labels=c("not living with both parents","living with both parents"))
```

parental level

```{r}
#level of education completed
#Fathers:
indrespb$paedu<-c(0)
indrespb$paedu[indrespb$ip_hiqual_dv%in%c(9)]<-1 #low (everything below)
indrespb$paedu[indrespb$ip_hiqual_dv%in%c(3,4,5)]<-2 #middle (completed A-level or secondary high school)
indrespb$paedu[indrespb$ip_hiqual_dv%in%c(1,2)]<-3 #high (have a degree)
indrespb$paedu[indrespb$paedu==0]<-NA

print(attr(indrespb$ip_hiqual_dv,"labels"))
indrespb$paedu<-factor(indrespb$paedu,levels=c(1,2,3),labels=c('low',"middle","high"))

#Mothers:
indrespb$maedu<-c(0)
indrespb$maedu[indrespb$im_hiqual_dv%in%c(9)]<-1 #low (everything below)
indrespb$maedu[indrespb$im_hiqual_dv%in%c(3,4,5)]<-2 #middle (completed A-level or secondary high school)
indrespb$maedu[indrespb$im_hiqual_dv%in%c(1,2)]<-3 #high (have a degree)
indrespb$maedu[indrespb$maedu==0]<-NA

indrespb$maedu<-factor(indrespb$maedu,levels=c(1,2,3),labels=c('low',"middle","high"))


#the highest education level between two parents
indrespb<-transform(indrespb, parentedu = pmin(im_hiqual_dv, ip_hiqual_dv))
indrespb$parentedu<-ifelse(is.na(indrespb$parentedu), indrespb$im_hiqual_dv, indrespb$parentedu)
indrespb$parentedu<-ifelse(is.na(indrespb$parentedu), indrespb$ip_hiqual_dv, indrespb$parentedu)

indrespb$parent_edu<-c(0)
indrespb$parent_edu[indrespb$parentedu%in%c(9)]<-1 #low (everything below, the category "other" is included)
indrespb$parent_edu[indrespb$parentedu%in%c(3,4,5)]<-2 #middle (complete A-level or have some qualifications)
indrespb$parent_edu[indrespb$parentedu%in%c(1,2)]<-3 #high (have a degree)
indrespb$parent_edu[indrespb$parentedu==0]<-NA

indrespb$parent_edu<-factor(indrespb$parent_edu,levels=c(1,2,3),labels=c('low',"middle","high"))

#two-category education
indrespb$parent_edu2<-c(0)
indrespb$parent_edu2[indrespb$parentedu%in%c(1,2)]<-1 #high (have a degree)
indrespb$parent_edu2<-factor(indrespb$parent_edu,levels=c(0,1),labels=c('have a degree',"do not have a degree"))

indrespb<-indrespb%>%dplyr::select(-parentedu)

#general health of parents**
indrespb<-indrespb%>%mutate(health_father=coalesce(ip_sf1, ip_scsf1))
indrespb<-indrespb%>%mutate(health_mother=coalesce(im_sf1, im_scsf1))
#least healthy parent
indrespb<-transform(indrespb, parenthealth = pmax(health_father, health_mother))
indrespb$parenthealth<-ifelse(is.na(indrespb$parenthealth), indrespb$health_father, indrespb$parenthealth)
indrespb$parenthealth<-ifelse(is.na(indrespb$parenthealth), indrespb$health_mother, indrespb$parenthealth)

#parent's age (youngest)
indrespb <- transform(indrespb, parentage = pmin(ip_age_dv
, im_age_dv))
indrespb$parentage<-ifelse(is.na(indrespb$parentage), indrespb$ip_age_dv, indrespb$parentage)
indrespb$parentage<-ifelse(is.na(indrespb$parentage), indrespb$im_age_dv, indrespb$parentage) 

#parent's income (sum and logged)
indrespb$palnincome<-indrespb$ip_fimnnet_dv+indrespb$im_fimnnet_dv
indrespb$palnincome[indrespb$palnincome<0]<-0
#add 1 to all income values to avoid -inf in log transformation
indrespb$paincomeadd<-indrespb$palnincome+1
indrespb$palnincome<-log(indrespb$paincomeadd)
indrespb<-indrespb%>%dplyr::select(-paincomeadd)

#parental tenure
indrespb$mateure<-c(0)
indrespb$mateure[indrespb$im_tenure_dv%in%c(1,2)]<-1 #ownership
indrespb$mateure[indrespb$im_tenure_dv%in%c(3,4,5)]<-2 #social rent
indrespb$mateure[indrespb$im_tenure_dv%in%c(6,7,8)]<-3 #private rent n other
indrespb$mateure[indrespb$mateure==0]<-NA

indrespb$mateure<-factor(indrespb$mateure,levels=c(1,2,3),labels=c("ownership","social rent","private rent n other"))

indrespb$pateure<-c(0)
indrespb$pateure[indrespb$ip_tenure_dv%in%c(1,2)]<-1 #ownership
indrespb$pateure[indrespb$ip_tenure_dv%in%c(3,4,5)]<-2 #social rent
indrespb$pateure[indrespb$ip_tenure_dv%in%c(6,7,8)]<-3 #private rent n other
indrespb$pateure[indrespb$pateure==0]<-NA

indrespb$pateure<-factor(indrespb$pateure,levels=c(1,2,3),labels=c("ownership","social rent","private rent n other"))

#the parents who have the most adv housing
indrespb <- transform(indrespb, housingtenure = pmin(ip_tenure_dv, im_tenure_dv))
indrespb$housingtenure<-ifelse(is.na(indrespb$housingtenure), indrespb$ip_tenure_dv, indrespb$housingtenure)
indrespb$housingtenure<-ifelse(is.na(indrespb$housingtenure), indrespb$im_tenure_dv, indrespb$housingtenure) 

indrespb$housing_tenure<-c(0)
indrespb$housing_tenure[indrespb$housingtenure%in%c(1,2)]<-1 #ownership
indrespb$housing_tenure[indrespb$housingtenure%in%c(3,4,5)]<-2 #social rent
indrespb$housing_tenure[indrespb$housingtenure%in%c(6,7,8)]<-3 #private rent n other
indrespb$housing_tenure[indrespb$housing_tenure==0]<-NA
indrespb$housing_tenure<-factor(indrespb$housing_tenure,levels=c(1,2,3),labels=c("ownership","social rent","private rent n other"))

indrespb<-indrespb%>%dplyr::select(-housingtenure)
```

```{r}
#cross_wave n interaction variables
#age * parental income
indrespb$paincomeXage<-indrespb$palnincome*indrespb$age_centered

#cross wave identifiers
indrespb$wave<-c(0)
indrespb$wave[indrespb$from_wave_h==1]<-1
indrespb$wave[indrespb$from_wave_i==1]<-2
indrespb$wave[indrespb$from_wave_j==1]<-3
indrespb$wave[indrespb$from_wave_k==1]<-4
indrespb$wave[indrespb$from_wave_l==1]<-5

indrespb$wave<-factor(indrespb$wave,levels=c(1,2,3,4,5),labels=c("wave 8","wave 9","wave 10","wave 11","wave 12"))

#indrespb<-indrespb%>%dplyr::select(-from_wave_h,-from_wave_i,-from_wave_j,-from_wave_k,-from_wave_l)

```

#### Depdendent variables

```{r}
indrespb$leave<-c(0)
indrespb$leave[indrespb$separated%in%c(1,2)]<-1

#wave h leavers
hleave<-indrespb[which(indrespb$leave==1&indrespb$from_wave_h==1),]
#join with i indresp for information
hleave<-join_all(list(hleave,i_indresp), by='pidp', type='left')

hleave$out<-c(0)
#leaving for marriage
hleave$out[hleave$i_mastat_dv%in%c(2,3)]<-1

#leaving for cohabitation
hleave$out[hleave$i_mastat_dv%in%c(10)]<-2

#leaving as a full-time student
hleave$out[hleave$i_jbstat%in%c(7)]<-3

#leaving for other reasons=0


#wave i leavers
ileave<-indrespb[which(indrespb$leave==1&indrespb$from_wave_i==1),]
#join with j indresp for information
ileave<-join_all(list(ileave, j_indresp), by='pidp', type='left')

ileave$out<-c(0)
#leaving for marriage
ileave$out[ileave$j_mastat_dv%in%c(2,3)]<-1

#leaving for coiabitation
ileave$out[ileave$j_mastat_dv%in%c(10)]<-2

#leaving as a full-time student
ileave$out[ileave$j_jbstat%in%c(7)]<-3

#leaving for other reasons=0


#wave j leavers
jleave<-indrespb[which(indrespb$leave==1&indrespb$from_wave_j==1),]
#join with k indresp for information
jleave<-join_all(list(jleave, k_indresp), by='pidp', type='left')

jleave$out<-c(0)
#leaving for marriage
jleave$out[jleave$k_mastat_dv%in%c(2,3)]<-1

#leaving for coiabitation
jleave$out[jleave$k_mastat_dv%in%c(10)]<-2

#leaving as a full-time student
jleave$out[jleave$k_jbstat%in%c(7)]<-3

#leaving for other reasons=0

#wave k leavers
kleave<-indrespb[which(indrespb$leave==1&indrespb$from_wave_k==1),]
#join with l indresp for information
kleave<-join_all(list(kleave, l_indresp), by='pidp', type='left')

kleave$out<-c(0)
#leaving for marriage
kleave$out[kleave$l_mastat_dv%in%c(2,3)]<-1

#leaving for coiabitation
kleave$out[kleave$l_mastat_dv%in%c(10)]<-2

#leaving as a full-time student
kleave$out[kleave$l_jbstat%in%c(7)]<-3

#leaving for other reasons=0

#wave l leavers
lleave<-indrespb[which(indrespb$leave==1&indrespb$from_wave_l==1),]
#join with m indresp for information
lleave<-join_all(list(lleave, m_indresp), by='pidp', type='left')

lleave$out<-c(0)
#leaving for marriage
lleave$out[lleave$m_mastat_dv%in%c(2,3)]<-1

#leaving for coiabitation
lleave$out[lleave$m_mastat_dv%in%c(10)]<-2

#leaving as a full-time student
lleave$out[lleave$m_jbstat%in%c(7)]<-3

#leaving for other reasons=0
out<-rbind.fill(hleave,ileave,jleave,kleave,lleave)
out<-out%>%dplyr::select(pidp,wave,out)
indrespb<-join_all(list(indrespb,out), by=c("pidp", "wave"), type='left')
```

```{r}
#mar
indrespb$mar<-c(0)
indrespb$mar[indrespb$out==1]<-1

#cohab
indrespb$cohab<-c(0)
indrespb$cohab[indrespb$out==2]<-1

#student
indrespb$student<-c(0)
indrespb$student[indrespb$out==3]<-1

#other
indrespb$other<-c(0)
indrespb$other[indrespb$out==0]<-1
```

## Model building

### Wave 1-5

### Wave 8-12

```{r}
indrespby<-indrespb[which(indrespb$age_centered%in%c(0:14)),]
indrespbfy<-indrespb[which(indrespb$age_centered%in%c(0:14)&indrespb$sex==1),]
indrespbmy<-indrespb[which(indrespb$age_centered%in%c(0:14)&indrespb$sex==0),]

mar_bf<-glm(mar~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+factor(liveboth)+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbfy,family=binomial(link="logit"))

cohab_bf<-glm(cohab~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+factor(liveboth)+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbfy,family=binomial(link="logit"))

student_bf<-glm(student~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+factor(liveboth)+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbfy,family=binomial(link="logit"))

all_bf<-glm(leave~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+factor(liveboth)+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbfy,family=binomial(link="logit"))
```

```{r}
#paincomeXage effects more strongly for men
mar_bm<-glm(mar~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbmy,family=binomial(link="logit"))


cohab_bm<-glm(cohab~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbmy,family=binomial(link="logit"))

student_bm<-glm(student~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbmy,family=binomial(link="logit"))

all_bm<-glm(leave~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespbmy,family=binomial(link="logit"))
```

```{r}
mar_b<-glm(mar~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+factor(sex)+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespby,family=binomial(link="logit"))


cohab_b<-glm(cohab~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+factor(sex)+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespby,family=binomial(link="logit"))

student_b<-glm(student~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+factor(sex)+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespby,family=binomial(link="logit"))

other_b<-glm(other~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+factor(sex)+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespby,family=binomial(link="logit"))

all_b<-glm(leave~palnincome+paincomeXage+factor(parent_edu)+factor(housing_tenure)+parenthealth+parentage+age_centered+factor(sex)+lnincome+factor(edu)+factor(family_status)+factor(racel2)+factor(student)+factor(unemployed)+factor(urban)+factor(houseprices)+factor(wave),data=indrespby,family=binomial(link="logit"))
```

parentincomeXage has a stronger effect on men concerning leaving the parental homes without destination specifications in wave 8-12 (leave)
